### 2019-05-23 fractional/CMJ style files/article-template.tex
seeing the light in the tunnel

\documentclass{article}
\usepackage{college-math-j}
\usepackage{xfrac}

%% IF YOU HAVE FONTS INSTALLED
%\usepackage{mtpro}
%\usepackage{mathtime}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{definition}
\newtheorem*{definition}{Definition}
\newtheorem*{remark}{Remark}

\begin{document}

\title{Visualization of [fulle name] Fractional Integrals}
\author{Woodrow Wilson and Herbert Hoover} %%leave blank in initial submission to allow for double blind reviewing
%\author{ }

\maketitle

%Notes authors should leave bios out altogether.

%\begin{biog} %comment out in initial submission to allow for double blind reviewing
%\item[\biogpic{\includegraphics[width=84pt]{Woodrow_Wilson.pdf}}Woodrow Wilson] (twwilson@princeton.edu) received his PhD in history and political science from Johns Hopkins University. He held visiting positions at Cornell and Wesleyan before joining the faculty at Princeton, where he was eventually appointed president of the university.  Among his proudest accomplishments was the abolition of eating clubs at Princeton on the grounds that they were elitist.

%\item[\biogpic{\includegraphics[width=84pt]{Herbert_Hoover.pdf}}Herbert Hoover] (hchoover@stanford.edu) entered Stanford University in 1891, after failing all of the entrance exams except mathematics.  He received his BS degree in geology in 1895, spent time as a mining engineer, then was appointed by his co-author to the U.S. Food Administration and the Supreme Economic Council, where he orchestrated the greatest famine relief efforts of all time.
%\end{biog}
%This is not to say that there  This is partly, because there is no simplistic geometric interpretation for fractional calculus.  
\noindent
Leibniz, who is regarded by many scholars as the farther of calculus, is credited with inventing the following notation: $\frac{d^n y}{d x^n}$. 
A natural question which arises when one inspects this notation is: Can $n$ be a non-integer value? L'Hospital phrased it differently: ``\emph{What if $n$ be a $\frac{1}{2}$?}''. 
In a letter Leibniz sent to L'Hospital, he himself commented on this intriguing question: ``\emph{This is an apparant paradox from which, one day, useful 
consequences will be drawn}.''

The mathematical discipline in which we study the extension of derivatives and integrals to non-integer orders is known as fractional calculus. Liouville and Riemann laid most 
of theoretical groundwork which was needed to develop the field of fractional calculus. Fractional calculus has now developed into a mature discipline. 
However, there is one aspect of the discipline which only started to develop fairly recently: its physical and geometric interpretability. At the first international conference on fractional calculus in
New Haven (USA) (which took place in 1974) it was pointed out that literature lacked an acceptable geometric and physical interpretation of fractional calculus. Subsequently, the last few decades has seen 
a number of papers that attempt to rectify this. There now exists probabilistic interpretations, geometric interpretations, physical interpretations and ecenomic 
interpretations of fractional order integrals and differentials.

In this article we specifically focus on the geometric interpretation of fractional integrals. To be more specific, we present a novel geometric interpretation of the Riemann-Liouville fractional integral.
As we mentioned earlier, Riemann and Liouville played an important role in the development of fractional calculus. Their work was crucial in defining what is now know as 
the Riemann-Lioville fractional integral. At this point in time the reader should be made aware of the fact that there exists many other accepted fractional integral definitions in the literature.
We, however, do not discuss them any further in this article. Investigating whether one can apply the proposed framework to other fractional integral definitions will be adressed in a 
future endeavour.

Moreover, the geometric interpretation we present here is especially useful from a pedigogical point of view as it is similar in nature to the definition of classical integration. The Riemann integral 
can be interpreted as an infinite sum of rectangular infinitesimals. In this paper, we show that the Riemann-Liouville fractional integral can be interpreted 
as an infinite sum of non-rectangular infinitesimals whose shape is determined by the order of integration $\alpha$ and the integration limit $x$.

The geometric interpretation we present in this paper is an extension of the fractional geometric framework presented by Podlubny. Podlubny showed that a fractional Riemann-Liouville 
integral can be converted to a Riemann-Stieltjes integral. Furthermore, Grobler showed that a Riemann-Stieltjes integral can be interpreted as an infiniesimal sum of 
non-rectangular infinitesimels, i.e. a Riemann-Stieltjes integral can be converted into a Cavalieri integral. If we combine these two ideas together we obtain the geometric interpretaion contained in the 
previous paragraph.

We specifically wrote this paper for people with a pedagogical interest, undergraduate students and for the layman who finds integration theory intriguing and 
as such we will try to steer clear from using too many mathematical definitions and proofs. We will rather convey our ideas using as many examples as possible. From the 
three groups mentioned we believe that the undergraduate student will find this article the most useful. Although undergraduate students have a limited understanding of 
fractional calculus they generally are quite interested in the subject. This is quite evident from perusing mathematical social media platforms and from surveying 
articles published in undergraduate journals []. One of the reasons undergraduates struggle to grasp fractional calculus, apart from it not normally being taught as part of the undergraduate curricilum, is because, 
it is such an abstract concept. We believe that the geometric interpretation we present in this paper will provide undergraduates with the insight that they need to quickly grasp and understand the fundamental 
concepts that unerpin fractional calculus. We also hope that this paper will enable and encourage undergraduate lectures to already introduce students to this fascinating mathematical discipline as part of their 
undergraduate training.

We start the paper by reviewing the definition of the Riemann-Stieltjes and Cavalieri integral. We then present the definition of the Riemann-Lioville fractional integral. 
We end the paper by presenting our novel geometric interpretation and a conclusion.

%Due to the 
%Even though most undergraduates have a limited understanding of fractional-order calculus they generally are quite interested in the subject. This is quite 
%evident from perusing mathematical social media platforms and from surveying articles published in undergraduate journals []. One of the reason fractional calculus is not generally 
%taught at the undegraduate level is due to the fact that the notion is quite abstract and it is therefore difficult to interpret geometrically. In this paper adress this shortcomming by presenting  
%a novel geometric interpretation of the Riemann-Liouville fractional integral which is similar in nature to the classical interpretation of classical integration. Our hope is that this 
%geometric interpretation will encourage and enable undergraduate lectures to introduce undergraduates to this fascinating mathematical discipline.

%Our hope is that the geometric 
%interpretation we present here will encourage and enable undergraduate lectures to introduce undergraduates to this fascinating mathematical discipline.

%We therefore only present the min-
%imum number of mathematical definitions that are needed to convey the new geometric framework.

%This article is not an in depth study into the field of integral calculus, but should rather be viewed as an introductory text

%It is important to point out at this point that there exists numerous acccepted definitions of fractional integration and differentiation. For the sake of 
%simplicity we decided to focus on one In this paper we will focus 
%on the Riemann-Lioville fractional integral.

%There are many accepted definitions of fractional order differentiation and integration. In this paper we will only focus

%In particular, their contribution played a crucial part in defining what is now known as the 
%Riemann-Lioville fractional integral. In this paper, we present a novel geometric interpretation of the Riemann-Lioville fractional integral which is similar 
%in nature to the classical interpretation of an ordinary Riemann integral. An ordinary Riemann integral can be interpreted as an infinitesimal sum of rectangular 
%infinitesimals.

%In this paper we present a novel geometric interpretation of the Riemann-Liouville fractional integral. There are in fact many accepted definitions of fractional integration 
%The interpretation we present here is especially useful from 
%a pedigogical point of view as it is similar in nature to the definition of classical integration. as an infinite sum of rectangular infinitesimals. In this paper, we demonstrate at the hand of examples that the Riemann-Liouville fractional integral can be interpreted 
%as an infinite sum of non-rectangular infinitesimals whose shape is determined by the order of integration $\alpha$ and the integration limit $x$. Furthermore, please note that the intended audience of this 
%paper is undergraduate students and lecturers and as such we do not present .  In ordinary integration we can interpret the integral

%Fractional calculus can be summarized as the generalization of integer-order differentiation and integration to an arbitrary order.
%Fractional-order calculus is as old as classical calculus itself. In this paper, we will mainly focus our attention on fractional-order integration and in particular investigate the Riemann-Liouville fractional 
%integral. It is Liouville and Riemann who laid much of the groundwork which was needed to define what is now known as the Riemann-Liouville fractional integral [Laurent’].

%The idea of integer-order differentiation and integration is well known to most undergraduates. However, undergraduates are less familiar with the idea of an arbitrary order differential or integral (i.e. fractional-order calculus).
%Even though most undergraduates have a limited understanding of fractional-order calculus they generally are quite interested in the subject. This is quite 
%evident from perusing mathematical social media platforms and from surveying articles published in undergraduate journals []. One of the reason fractional calculus is not generally 
%taught at the undegraduate level is due to the fact that the notion is quite abstract and it is therefore difficult to interpret geometrically. In this paper adress this shortcomming by presenting  
%a novel geometric interpretation of the Riemann-Liouville fractional integral which is similar in nature to the classical interpretation of classical integration. Our hope is that this 
%geometric interpretation will encourage and enable undergraduate lectures to introduce undergraduates to this fascinating mathematical discipline.

\section{Riemann-Stieltjes Integral}
Recall that the Riemann-Stieltjes integral is defined as follows:
\begin{equation}
\label{eq:g_int2}
\int_{a'}^{b'} f(x) dg(x) =  \lim_{n \rightarrow \infty}\sum_{i=0}^{n-1} f(x_i^2)[g(x_{i+1}^2)-g(x_{i}^2)], 
\end{equation}
%Fig.~\ref{fig:2d_geo} depicts a region whose area can be determined by considering non--rectangular integration strips.
%. Note that the shape of the integration strips are determined by $a(y)$. 
%In Fig.~\ref{fig:2d_geo}, the function $a(y)$ is depicted in red. If we translate $a(y)$ along the $x$-axis we actually create two partitions on the $x$-axis, 
%i.e. $(x_i^1)_{i=0}^{n}$ and $(x_i^2)_{i=0}^{n}$ with $x_i^1\in[a,b]$ and $x_i^2\in[a',b']$.  To map partition points in $[a',b']$ to partition points 
%in $[a,b]$ we use the function $g$. The reverse is achieved through $h$. The functions $g$ and $h$ depend soly on $a(y)$ and are also 
%computed with it.
\noindent
with $(x_i^2)_{i=0}^n$ being arbitrary partition points on the $x$-axis. 
The function $g$ is known as the integrator and if it is monotone increasing then it maps points in the interval $[a',b,]$ to the interval $[a,b]$, with $a = g(a')$ and $b = g(b')$.
The integral in Eq.~\eqref{eq:g_int2} is hard to evaluate in its present form. If $g$ is differentiable, however, then it becomes trivial to 
evaluate a Riemann-Stieltjes integral due to the following identity:
\begin{equation}
\int_{a'}^{b'} f(x) dg(x) = \int_{a'}^{b'} f(x)g'(x)~dx.
\end{equation}
As an example:
\begin{equation}
\int_{\frac{1}{2}}^{2} f(x) dg(x)=\int_{\frac{1}{2}}^2 x~d2x = \int_{\frac{1}{2}}^2 2x~dx = 3.75. 
\end{equation}

\section{Cavalieri Integral}
Let the region $R$ be bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$ and $b(y)=4-y$. This region is shown in Fig.~\ref{fig:caval2}.\\
%\begin{figure}[htb]
%\centering
%\includegraphics[width=0.3\textwidth]{fig12}
%\caption{Region bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$, and $b(y)=4-y$. Reproduced from Quaestiones Mathematicae (2012) 35: 265-296 with permission \copyright~ NISC (Pty) Ltd.}
%\label{fig:ex1}
%\end{figure}
\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{fig13.eps}
\caption{Region bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$, and $b(y)=4-y$. The figure also depicts the partition points $x_i^2$ as used in the Cavalieri sum (see Eq.~\ref{eq:cav_sum}). Reproduced from Quaestiones Mathematicae (2012) 35: 265-296 with permission \copyright~ NISC (Pty) Ltd.}
\label{fig:caval2}
\end{figure}

\noindent
The area of $R$ can be determined as follows if we employ the classical notion of integration:
\begin{equation}
\int_0^2x\, dx+\int_2^44-x\, dx- \int_0^{\frac{1}{2}}x\, dx-\int_{\frac{1}{2}}^11-x\, dx = 3.75. 
\end{equation}

\noindent
There, however, exists a more straightforward way do determine the area of $R$, i.e. we can sum together the area of non-rectangular integration strips inscribing $R$ instead of
having to sum together the area of multiple rectangular integration strips. Moreover, the shape of the non-rectangular integration strips we should use is determined by the function $a(y)$. We can express this notion more formally as 
follows, let $(x_i^1)_{i=0}^{n}$ denote a partition on the $x$-axis, such that $a = x_0^1 < x_1^1 < \cdots < x_n^1 = b$, and $\Delta x_i^1 = x_{i+1}^1 - x_i^1$.
We are now able to construct the following sum (the lower Cavalieri sum):
\begin{equation}
\label{eq:cav_sum}
\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1.
\end{equation}
The partition points $(x_i^2)_{i=0}^{n}$ are depicted in Fig.~\ref{fig:caval2}. Note that Eq.~\eqref{eq:cav_sum} approximates the area of $R$. In the limit Eq.~\eqref{eq:cav_sum} approaches 
the Cavalieri integral:
\begin{equation}
\label{eq:caval1}
\int_{a(y)}^{b(y)}f(x)\, dx = \lim_{n\to \infty}\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1.
\end{equation}
It is, however, quite hard to evaluate the above integral directly. Fortunately, it is easy to convert a Cavalieri integral into an equivalent Riemann or Riemann-Stieltjes 
integral by using the transformation functions $h$ and $g$. Expressed mathematically:
\begin{equation}
\label{eq:main_cav}
\int_{a(y)}^{b(y)}f(x)\,dx =\int_a^b f \circ h (x)\, dx = \int_{a'}^{b'} f(x) dg(x).
\end{equation}
We can calculate $g$ as follows:
\begin{equation}
g(x) = x - a\circ f(x) + a.
\end{equation}
Moreover, $h=g^{-1}$. We can now evaluate the Cavalieri integral by using its equivalent Riemann or Rieman-Stieltjes integral.
If we use its Riemann equivalent we obtain:
\begin{equation}
\int_{a(y)}^{b(y)}f(x)\, dx = \int_a^b f \circ h (x)\, dx = \dfrac{1}{2}\int_1^4x\, dx = 3.75.  
\end{equation}
If we use its Riemann-Stieltjes equivalent we obtain:
\begin{equation}
\int_{a(y)}^{b(y)}f(x)\, dx = \int_{a'}^{b'} f \, dg(x) = \int_{\frac{1}{2}}^2x\, d2x = 3.75.  
\end{equation}

\noindent
Conversely, if $g$ is known (and not $a(y)$) and $g(a') = a$ then we can calculate $a(y)$ using the following:
\begin{equation}
\label{eq:a_y}
a(y) = f^{-1}(y) - g\circ f^{-1}(y) + g(a'). 
\end{equation}
Note the power of Eq.~\eqref{eq:a_y} it allows us to transform a Riemann-Stieltjes integral into an equivalent Cavalieri integral. 
Furthermore, this transformation enables us to assign a geometric interpretation to a Riemann-Stieltjes integral. The interpretation being:
it represents the area obtained by summing together the area of an infinite number of non-rectangular infinitesimals whose shape is determined by $a(y)$.

\section{Gamma Function}
The gamma function was first devised by Euler. The gamma function is defined for all real numbers except for the non-negative integers. Moreover, it is a natural extension of the factorial function to the real numbers. The Gamma function is defined as:
\begin{equation}
\Gamma(x) = \int_0^{\infty} t^{x-1} e^{-t}\,dt.
\end{equation}
The function $\Gamma(x)$ is depicted in Fig.~\ref{fig:gamma} for $x\in(-5,5)$.
\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{gamma.pdf}
\caption{A plot of $\Gamma(x)$ for $x\in(-5,5)$. $\Gamma(x)$ is not defined for the non-negative integers.}
\label{fig:gamma}
\end{figure}

\section{Riemann-Liouville Fractional Integral}
In this section we present the definition of the Riemann-Liouville fractional integral of order $\alpha$. Let
\begin{equation}
If(x) := \int_0^x f(t) dt.
\end{equation}
If we apply the above operator in a repetitive manner to $f$ we obtain the $n$th antiderivative of $f$ based at 0:
\begin{equation}
\label{eq:n_anti}
I^nf(x) = \int_0^x\int_0^{x_1}\cdots \int_0^{x_{n-1}}f(x_n)~dx_n\cdots dx_2 dx_1.
\end{equation}
Applying Cauchy's formula for repetitive integration to the above $n$th antiderivative of $f$ helps us express the above $n$th antiderivative as a single integral. 
Applying Cauchy's formula to Eq.~\eqref{eq:n_anti} results in:
\begin{equation}
\label{eq:cauchy}
I^nf(x) = \frac{1}{(n-1)!}\int_0^x (x-t)^{n-1}f(t)~dt.
\end{equation}
As an example if $f(x)=x$ and $n=2$ then Eq.~\eqref{eq:cauchy} reduces to:
\begin{equation}
I^2f(x) = \int_0^x (x-t)t~dt = \frac{t^2x}{2} - \frac{t^3}{3} \Bigg |_0^x = \frac{x^3}{3!}.
\end{equation}
The definition in Eq.~\eqref{eq:cauchy} can now be extended to an arbitrary fractional order by replacing $(n-1)!$ with the Gamma function (see Fig.~\ref{fig:gamma}):
\begin{equation}
\label{eq:frac_integral}
I^{\alpha}f(x) = \frac{1}{\Gamma(\alpha)}\int_0^x (x-t)^{\alpha-1}f(t)~dt.
\end{equation}
Eq.~\eqref{eq:frac_integral} is known as the Riemann-Liouville fractional integral of order $\alpha$. As an example if $f(x)=x$ and $\alpha=\frac{1}{2}$ then 
Eq.~\eqref{eq:frac_integral} reduces to:
\begin{equation}
I^{\frac{1}{2}}f(x) = \frac{1}{\Gamma(\frac{1}{2})} \int_0^x \frac{t}{\sqrt{x-t}} dt = \frac{1}{\sqrt{\pi}}\left [ -\frac{2}{3}\sqrt{x-t}(t+2x)\right] \Bigg |_0^x=\frac{4}{3\sqrt{\pi}}x^{\frac{3}{2}}. 
\end{equation}
Furthermore, it is straightforward to show that the $I$ operator satisfies:
\begin{equation}
I^{\alpha}I^{\beta}f(x) = I^{\alpha+\beta}f(x). 
\end{equation}

\section{Geometric Interpretation of the Riemann-Liouville Fractional Integral}

Recall that the Riemann-Liouville fractional integral of order $\alpha$ is defined as:
\begin{equation}
\label{eq:RL_frac_int}
\frac{1}{\Gamma(\alpha)}\int_{0}^{t}f(\tau)(t-\tau)^{\alpha-1}~d\tau
\end{equation}

\noindent
Podlubny et al. showed that the above fractional integral can be rewritten as a Riemann-Stieltjes integral:
\begin{equation}
\label{eq:rs}
\int_0^{t} f(\tau)~dg_t(\tau),
\end{equation}
with 
\begin{equation}
\label{eq:g_rl}
g_t^{\alpha}(\tau) = \frac{\left \{t^{\alpha} - (t-\tau)^{\alpha} \right \}}{\Gamma(\alpha+1)}. 
\end{equation}

\noindent
Note that $g_t^{\alpha}(\tau)$ is always invertible, since $\frac{d g_t^{\alpha}}{d \tau} = \frac{(t-\tau)^{\alpha-1}}{\Gamma(\alpha)}>0$ for $\tau\in[0,t]$. The inverse 
of $g_t^{\alpha}(\tau)$ is equal to:
\begin{equation}
h_t^{\alpha}(\tau) = t - \sqrt[\alpha]{t^{\alpha} - \Gamma(\alpha+1)\tau}
\end{equation}

\noindent
Let $\mathsf{A} = \{0,0.2,0.4,0.6,0.8,1\}$. The analytic expressions of $g_{10}^{\alpha}(\tau)$ and $h_{10}^{\alpha}(\tau)$ are presented in Table~\ref{tab:gandh} for $\alpha\in\mathsf{A}$ and $t=10$. 
The expressions in Table~\ref{tab:gandh} are depicted in Fig~\ref{fig:gandh}.

\begin{table}[h!]
 \centering
 \caption{The analytic expressions of $g_{10}^{\alpha}(\tau)$ and $h_{10}^{\alpha}(\tau)$ for $\alpha\in\mathsf{A}$.}
 \label{tab:gandh}
 \begin{tabular}{|c || c | c|} 
 \hline
 $\alpha$ & $g_{10}^{\alpha}(\tau)$ & $h_{10}^{\alpha}(\tau)$ \\ [0.5ex] 
 \hline\hline
 $0$ & 0 & $y=0$ \\ 
 $0.2$ & $[\Gamma\left (\frac{6}{5} \right )]^{-1}\left(\sqrt[5]{10}-\sqrt[5]{10-\tau}\right)$ & $10 - \left ( \sqrt[5]{10} -  \Gamma\left (\frac{6}{5} \right ) \tau \right )^5$  \\
 $0.4$ & $[\Gamma\left (\frac{7}{5} \right )]^{-1}\left(\sqrt[5]{10^2}-\sqrt[5]{(10-\tau)^2}\right)$ & $10 - \sqrt{\left ( \sqrt[5]{10} -  \Gamma\left (\frac{7}{5} \right ) \tau \right )^5}$ \\
 $0.6$ & $[\Gamma\left (\frac{8}{5} \right )]^{-1}\left(\sqrt[5]{10^3}-\sqrt[5]{(10-\tau)^3}\right)$ & $10 - \sqrt[3]{\left ( \sqrt[5]{10} -  \Gamma\left (\frac{8}{5} \right ) \tau \right )^5}$ \\
 $0.8$ & $[\Gamma\left (\frac{9}{5} \right )]^{-1}\left(\sqrt[5]{10^4}-\sqrt[5]{(10-\tau)^4}\right)$ & $10 - \sqrt[4]{\left ( \sqrt[5]{10} -  \Gamma\left (\frac{9}{5} \right ) \tau \right )^5}$ \\ [1ex] 
 $1$ & $\tau$ & $\tau$ \\ [1ex] 
 \hline
 \end{tabular}
 \end{table}

\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{gh.pdf}
\caption{This figure depicts the curves of $g_{10}^{\alpha}(\tau)$ and $h_{10}^{\alpha}(\tau)$ for $\alpha\in\mathsf{A}$. The curves associated with $g$ are plotted using 
solid lines, while the curves plotted with dashed lines are associated with $h$. The lighter the shade with which a curve is depicted, the larger the $\alpha$-value is, that is associated with it.
}
\label{fig:gandh}
\end{figure}

\noindent
Furthermore, Grobler et. al showed that it is trivial to convert a Riemann-Stieltjes integral into a Cavalieri integral. We review this conversion procedure in the Cavalieri section of this paper (see Fig.~\ref{fig:caval2}). If we apply their conversion steps to Eq.~\eqref{eq:rs} we obtain:
\begin{equation}
\label{eq:cav}
\int_{a_t^{\alpha}(y)}^{b_t^{\alpha}(y)} f(\tau)~d\tau, 
\end{equation}
with
\begin{equation}
\label{eq:cav_a}
a_t^{\alpha}(y) = f^{-1}(y) - \frac{t^{\alpha}-(t-f^{-1}(y))^{\alpha}}{\Gamma(\alpha+1)}.
\end{equation}
Eq.~\eqref{eq:cav_a} is obtained by substituting Eq.~\eqref{eq:g_rl} into Eq.~\eqref{eq:a_y}.

\noindent
We can easily assign a geometric interpretation to Eq.~\eqref{eq:cav}, since it is a Cavalieri integral (see Fig.~\ref{fig:caval2}). The fractional integral in Eq.~\eqref{eq:RL_frac_int} can, therefore, be interpreted as the area obtained 
by summing together the area of an infinite number of infinitesimally small non-rectangular integration strips whose shape is determined by $\alpha$ and $t$. If $\alpha$ is equal to one, however, then the integral reduces to a normal Riemann integral as the integration strips become rectangular for this particular choice of $\alpha$ (its shape becomes independent of $t$).\\

\noindent
Interestingly, Eq.~\eqref{eq:main_cav} implies that we can use $h_t^{\alpha}(\tau)$ to convert Eq.~\eqref{eq:cav} into the following Riemann integral:
\begin{equation}
\label{eq:option2}
\int_0^{\frac{t^{\alpha}}{\Gamma(\alpha+1)}} f\circ h_{\alpha}^t (\tau) d\tau.  
\end{equation}
We can, therefore, evaluate Eq.~\eqref{eq:cav} in one of two ways: we can either evaluate Eq.~\eqref{eq:rs} or we can evaluate Eq.~\eqref{eq:option2}.

% \begin{table}[h!]
% \centering
%  \begin{tabular}{|c || c | c|} 
%  \hline
%  & \multicolumn{2}{|c|}{$a_{10}^{\alpha}(\tau)$}\\
%  \hline
%  &$f(\tau) = \tau$ & $f(\tau)=\sqrt{\tau}$ \\ [0.5ex] 
%  \hline\hline
%  $0$ & 0 & $y=0$ \\ 
%  $0.2$ & $y-[\Gamma\left (\frac{6}{5} \right )]^{-1}\left(\sqrt[5]{10}-\sqrt[5]{10-y}\right)$ & $y^2-[\Gamma\left (\frac{6}{5} \right )]^{-1}\left(\sqrt[5]{10}-\sqrt[5]{10-y^2}\right)$  \\
%  $0.4$ & $y-[\Gamma\left (\frac{7}{5} \right )]^{-1}\left(\sqrt[5]{10^2}-\sqrt[5]{(10-y)^2}\right)$ & $y^2-[\Gamma\left (\frac{7}{5} \right )]^{-1}\left(\sqrt[5]{10^2}-\sqrt[5]{(10-y^2)^2}\right)$ \\
%  $0.6$ & $y-[\Gamma\left (\frac{8}{5} \right )]^{-1}\left(\sqrt[5]{10^3}-\sqrt[5]{(10-y)^3}\right)$ & $y^2-[\Gamma\left (\frac{8}{5} \right )]^{-1}\left(\sqrt[5]{10^3}-\sqrt[5]{(10-y^2)^3}\right)$ \\
%  $0.8$ & $y-[\Gamma\left (\frac{9}{5} \right )]^{-1}\left(\sqrt[5]{10^4}-\sqrt[5]{(10-y)^4}\right)$ & $y^2-[\Gamma\left (\frac{9}{5} \right )]^{-1}\left(\sqrt[5]{10^4}-\sqrt[5]{(10-y^2)^4}\right)$ \\ [1ex] 
%  $1$ & $\tau$ & $\tau$ \\ [1ex] 
%  \hline
%  \end{tabular}
% \end{table}

% \begin{table}[h!]
% \centering
%  \begin{tabular}{|c || c|} 
%  \hline
%  Symbol & Definition  \\ [0.1ex]
%  \hline 
%  \rule{0pt}{2.5ex}
%  $\mathsf{G}^{10}$ & $\{g_{10}^{\alpha}(\tau)|\alpha\in\{0,\frac{1}{5},\frac{2}{5},\frac{3}{5},\frac{4}{5}\}\}$\\ [0.5ex]
%  $\mathsf{H}^{10}$ & $\{h_{10}^{\alpha}(\tau)|\alpha\in\{0,\frac{1}{5},\frac{2}{5},\frac{3}{5},\frac{4}{5}\}\}$\\ [0.5ex]
%  $\mathsf{A}^{10}$ & $\{a_{10}^{\alpha}(\tau)|\alpha\in\{0,\frac{1}{5},\frac{2}{5},\frac{3}{5},\frac{4}{5}\}\}$\\ [0.5ex]
%  $\mathsf{B}^{10}$ & $\{b_{10}^{\alpha}(\tau)|\alpha\in\{0,\frac{1}{5},\frac{2}{5},\frac{3}{5},\frac{4}{5}\}\}$\\ [0.5ex]
%  $\mathsf{B}^{\frac{4}{5}}_1$ & $\{b_{t}^{\frac{4}{5}}(\tau)|t\in\alpha\in\{2,4,6,8,10\}$\\ [0.5ex]
%  $\mathsf{B}^{\frac{4}{5}}_2$ & $\{b_{10}^{\frac{4}{5}}(\tau)-c|c\in\in\{0,2,4,6,8\}$\\ [0.5ex]
%  \hline 
%  \end{tabular}
% \end{table}

\section{Examples}
In this section we will illustrate the usefulness of the non-static non-rectangular sum-based geometric interpretation we presented in the previous section at the 
hand of two examples.\\

\noindent
Let us first consider the following fractional integral:
\begin{equation}
\label{eq:ex1_raw}
\frac{1}{\Gamma(\alpha)}\int_0^t f(\tau) (t-\tau)^{\alpha-1}~d\tau = \frac{1}{\Gamma(\alpha)}\int_0^t \tau(t-\tau)^{\alpha-1}~d\tau. 
\end{equation}
The function $f(\tau) =\tau$ is depicted in Fig.~\ref{fig:geo1}. As we mentioned in the previous section, we can assign a geometric interpretation to the above fractional integral by converting it into a Cavalieri integral using two steps. First, we convert 
the above integral into a Riemann-Stieltjes integral:
\begin{equation}
\frac{1}{\Gamma(\alpha)}\int_0^t f(\tau) (t-\tau)^{\alpha-1}~d\tau = \int_0^t f(\tau)~dg_t^{\alpha}(\tau). 
\end{equation}
We then convert the resulting Riemann-Stieltjes integral into a Cavalieri integral:
\begin{equation}
\label{eq:ex1}
\frac{1}{\Gamma(\alpha)}\int_0^t f(\tau) (t-\tau)^{\alpha-1}~d\tau = \int_0^t f(\tau)~dg_t^{\alpha}(\tau)=\int_{a_t^{\alpha}(y)}^{b_t^{\alpha}(y)} f(\tau)~d\tau. 
\end{equation}

\begin{figure}[htb]
\centering
\includegraphics[width=1\textwidth]{geo_int1.pdf}
\caption{The non-static non-rectangular sum-based geometric interpretation of Eq.~\eqref{eq:ex1_raw}.}
\label{fig:geo1}
\end{figure}

\noindent
Since Eq.~\ref{eq:ex1}, is a Cavalieri integral it has a known geometric interpretation (see Fig.~\ref{fig:caval2}). It can be interpreted as the area obtained by summing together the areas of an infinite 
number of infinitesimals whose shape is determined by $\alpha$ and $t$. Moreover, these infinitesimals have to be translations of $b_t^{\alpha}(y)$ which, in turn, is 
function of $\alpha$ and $t$. The geometric interpretation of Eq.~\eqref{eq:ex1_raw} is, therefore, non-static as it depends on the value of $\alpha$ and $t$. 
There is one thing, however, which remains unclear: in what way does $\alpha$ and $t$ affect the shape of the aforementioned infinitesimals, i.e. how does the value of $\alpha$ and $t$ affect the shape of $b_t^{\alpha}(y)$. 
We will answer this question by using two basic steps. First off, we will fix the value of $\alpha$ and $t$ in Eq.~\eqref{eq:ex1} and then depict $b_t^{\alpha}(y)$. We will then explore how varying the value of $\alpha$ and $t$ in Eq.~\eqref{eq:ex1} affects the shape $b_t^{\alpha}(y)$.\\

%We know that a Cavalieri integral can be interpreted as the area which is obtained by summing together the area associated with an infinite number of non-rectangular infinitesimals whose shape 
%is determined by the integration limits of the Cavalieri integral. Note that neither $g_t^{\alpha}(\tau)$ nor $a_t^{\alpha}(y)$ are static functions, 
%they depend on the value of $\alpha$ and $t$. This implies that a fractional integral of order $\alpha$ and extent $t$ can be interpeted as the area obtained by summing together the area associated with an infinite number of infinitesimals whose 
%shape is determined by $\alpha$ and $t$, i.e. the geometric interpretation of Eq.~\eqref{eq:ex1_raw} is non-static as it depends on the value of $\alpha$ and $t$.

\noindent
If we set the value of $\alpha$ to $\sfrac{4}{5}$ and the value of $t$ to 10 in Eq.~\eqref{eq:ex1} we obtain:
\begin{equation}
\label{eq:ex1_specific}
A_{10}^{\sfrac{4}{5}}=\int_{a_{10}^{\frac{4}{5}}(y)}^{b_{10}^{\frac{4}{5}}(y)} f(\tau)~d\tau 
\end{equation}
Let $R_{10}^{\sfrac{4}{5}}$ denote the region which is bounded by the $\tau$-axis, the function $f(\tau)$ and the function $b_{10}^{\sfrac{4}{5}}(y)$. The region $R_{10}^{\sfrac{4}{5}}$ is depicted in the right-most panel of Fig.~\ref{fig:geo1}. It is the largest of all the 
regions that are depicted in the right-most panel of Fig.~\ref{fig:geo1}. Moreover, let $A_{10}^{\sfrac{4}{5}}$ denote the area of $R_{10}^{\sfrac{4}{5}}$. For the sake of completeness, lets first verify that we can employ the Cavalieri integral in Eq.~\ref{eq:ex1_specific} to compute $A_{10}^{\sfrac{4}{5}}$ from first principles.
We can approximate the area of $R_{10}^{\sfrac{4}{5}}$ by summing together the area of $n$ equal width non-rectangular integration strips that inscribe $R_{10}^{\sfrac{4}{5}}$, i.e. we can 
construct a Cavalieri sum (see Eq.~\eqref{eq:cav_sum}). Note, the sides of the non-rectangular integration strips we use in this approximation have to be translations of $b_{10}^{\sfrac{4}{5}}(y)$.
If we take the limit of the aforementioned sum as $n\rightarrow \infty$ we approach $A_{10}^{\sfrac{4}{5}}$, i.e. the result obtained via the aforementioned limit is the Cavalieri integral found in Eq.~\eqref{eq:ex1_specific}. In conclusion, we can think of Eq.~\eqref{eq:ex1_specific} as implying that the region $R_{10}^{\sfrac{4}{5}}$ is made up 
of an infinite number of non-rectangular infinitesimals; all of them being translations of $b_{10}^{\sfrac{4}{5}}(y)$. Note, the right-most curve depicted in the right-most panel of Fig.~\ref{fig:geo1} is $b_{10}^{\sfrac{4}{5}}(y)$.  Moreover, the special case in which we approximate $A_{10}^{\sfrac{4}{5}}$ using five equal width non-rectangular integration strips is depicted in the right-most panel of Fig.~\ref{fig:geo1}. 
The sides of the aforementioned five integration strips are depicted in Fig.~\ref{fig:geo1} using dashed lines. Let, $\mathcal{B}_{\mathsf{T}}^{\sfrac{4}{5}}=\{b_{10}^{\frac{4}{5}}(y)-c|c\in(\mathsf{T}-2)\}$ with $\mathsf{T}=\{2,4,6,8,10\}$.\\

%Let $\mathsf{B}_{10}^{\sfrac{4}{5}}$ denote the set containing these dashed curves.
%We can, therefore, interpet Eq.~\eqref{}  The region $R_{10}^{\sfrac{4}{5}}$ is, therefore, made up 
%of an infinite number of non-rectangular infinitesimals; all of them being translations of $b_{10}^{\sfrac{4}{5}}(y)$.

\noindent
Varying both $\alpha$ and $t$ results in Eq.~\eqref{eq:ex1}. In contrast with Eq.~\eqref{eq:ex1_specific}, Eq.~\eqref{eq:ex1} can be associated with more than one region. 
Let $R_{t}^{\alpha}$ denote the region which is bounded by the $\tau$-axis, the function $f(\tau)$ and the function $b_{t}^{\alpha}(y)$. Note that $b_{t}^{\alpha}(y)$ forms the 
right-most edge of $R_{t}^{\alpha}$. Moreover, let $A_{t}^{\alpha}$ denote the area of $R_{t}^{\alpha}$ and let $\mathsf{R}_{\mathsf{T}}^{\mathsf{A}}=\{R_t^{\alpha}|t\in\mathsf{T},~\alpha\in\mathsf{A}\}$ and 
$\mathsf{B}_{\mathsf{T}}^{\mathsf{A}}=\{b_t^{\alpha}(y)|t\in\mathsf{T},~\alpha\in\mathsf{A}\}$. 
The regions in the set $\mathsf{R}_{\mathsf{T}}^{\mathsf{A}}$ are depicted in Fig.~\ref{fig:geo1}. The regions that are depicted in each of the panels of Fig.~\ref{fig:geo1} were generated by varying $t$, whilst keeping the value of $\alpha$ fixed. 
The title of each panel indicate which $\alpha$-value was used to generate the regions depicted in each panel. The only difference between the regions located in each 
panel is their right edges. The right edges of the regions in each panel are depicted using varying shades of gray. The lighter the shade with which an edge is depicted, 
the larger the $t$-value is, that is associated with it.\\

\noindent
As was the case for Eq.~\eqref{eq:ex1_specific} and $R_{10}^{\sfrac{4}{5}}$, Eq.~\eqref{eq:ex1} implies that we can assume that each of the regions depicted in Fig.~\ref{fig:geo1} are made up of an infinite number of infinitesimals. The infinitesimals that make up a specific region are translations of that region's right edge. 
The following observations can be made by inspecting Fig.~\ref{fig:geo1}: 
\begin{description}
 \item[Effect of $\alpha$] the grey-tinted solid curves depicted in the different panels of Fig.~\ref{fig:geo1} are not equal to one-another, i.e. $\mathsf{B}_{\mathsf{T}}^{\alpha_1}\neq\mathsf{B}_{\mathsf{T}}^{\alpha_2}$. 
 \item[Effect of $t$] the grey-tinted solid curves depicted in each of the panels of Fig.~\ref{fig:geo1} are not translations of one-another. Moreover, the dashed grey curves and the solid grey-tinted curves in the right-most panel are not equal to one-another, i.e. $\mathsf{B}_{\mathsf{T}}^{\sfrac{4}{5}}\neq\mathcal{B}_{\mathsf{T}}^{\sfrac{4}{5}}$.
\end{description}
We can therefore draw the following conclusion: the value of both $\alpha$ and $t$ alters the shape of the infinitesimals (i.e. the upper limit $b_{t}^{\alpha}(y)$ of Eq.~\eqref{eq:ex1} depends on $\alpha$ and $t$) that make up the regions in Fig.~\ref{fig:geo1}. Interestingly, the effect of $\alpha$ on the shape of the infinitesimals that make up the regions depicted in Fig.~\ref{fig:geo1} is larger than the effect that $t$ has on their shape. 
The curves associated with the infinitesimals obtained by varying $t$, whilst keeping the value of $\alpha$ fixed are very similar in nature and there exists a high degree of correlation between them (if their 
sizes are ignored). The curves associated with the infinitesimals obtained by varying $\alpha$, whilst keeping the value of $t$ fixed are quite different from one another (the aforementioned curves are not highly correlated with 
one another).\\

%In the case of $\alpha$: the grey-tinted solid curves depicted in each panel are not equal to one-another. Similarly, in the case of $t$: the dashed grey curves and the solid grey-tinted curves in the right-most panel are not equal to one-another, i.e.$\mathsf{B}_T^{\alpha}\neq\mathcal{B}_T^{\alpha}$.

\begin{table}[h!]
\centering
\caption{Curves obtained by evaluating Eq.~\eqref{eq:ex1} and Eq.~\eqref{eq:ex2} for all $\alpha\in \mathsf{A}$ using either Eq.~\eqref{eq:rs} or Eq.~\eqref{eq:option2}.}
\label{tab:eval}
\begin{tabular}{||c|| c| c||} 
 \hline
 $\alpha$ & $\frac{1}{\Gamma(\alpha)}\int_0^t \tau(t-\tau)^{\alpha-1}~d\tau$ & $\frac{1}{\Gamma(\alpha)}\int_0^t \sqrt{\tau}(t-\tau)^{\alpha-1}~d\tau$ \\[0.5ex]
 \hline\hline
 \rule{0pt}{2.5ex}
 $0$ & $t$ & $\sqrt{t}$  \\  
 \rule{0pt}{2.5ex}
 $0.2$ & $\frac{25}{6\Gamma(\frac{1}{5})} t^{\frac{6}{5}}$  & $\frac{\sqrt{\pi}}{2\Gamma(\frac{17}{10})} t^{\frac{7}{10}}$  \\ 
 \rule{0pt}{2.5ex}
 $0.4$ & $\frac{25}{14\Gamma(\frac{2}{5})} t^{\frac{7}{5}}$  & $\frac{\sqrt{\pi}}{2\Gamma(\frac{19}{10})} t^{\frac{9}{10}}$  \\
 \rule{0pt}{2.5ex}
 $0.6$ & $\frac{25}{24\Gamma(\frac{3}{5})} t^{\frac{8}{5}}$  & $\frac{\sqrt{\pi}}{2\Gamma(\frac{21}{10})} t^{\frac{11}{10}}$  \\ 
 \rule{0pt}{2.5ex}
 $0.8$ & $\frac{25}{36\Gamma(\frac{4}{5})} t^{\frac{9}{5}}$  & $\frac{\sqrt{\pi}}{2\Gamma(\frac{23}{10})} t^{\frac{13}{10}}$  \\ 
 \rule{0pt}{2.5ex}
 $1$ & $\frac{1}{2}t^2$  & $\frac{2}{3}t^{\frac{3}{2}}$  \\ [1ex]
 \hline
 \end{tabular}
\end{table}

\begin{figure}[htb]
\centering
\includegraphics[width=1\textwidth]{func_eval1.pdf}
\caption{Depicts the curves which are obtained by evaluating Eq.~\eqref{eq:ex1} for all $\alpha\in \mathsf{A}$ using either Eq.~\eqref{eq:rs} or Eq.~\eqref{eq:option2}.}
\label{fig:eval1}
\end{figure}

\noindent
If we evaluate the integral in Eq.~\eqref{eq:ex1} for all $\alpha\in \mathsf{A}$ (using either Eq.~\eqref{eq:rs} or Eq.~\eqref{eq:option2}) we obtain the curves in the 
left-most column of Table~\ref{tab:eval}. These curves are depicted in Fig.~\ref{fig:eval1}.
The area of the regions depicted in Fig.~\ref{fig:geo1} are also plotted in Fig.~\ref{fig:eval1} using circular markers. As expected, the circular markers in Fig.~\ref{fig:eval1} that correspond to the areas of the regions in Fig.~\ref{fig:eval1}  
associated with a particular choice of $\alpha$ (i.e. regions depicted in a particular panel of Fig.~\ref{fig:geo1}) fall on the curve in Fig.~\ref{fig:eval1}; associated with the same choice of $\alpha$. Moreover, note that we have adopted the 
same coloring scheme in Fig.~\ref{fig:geo1} and Fig.~\ref{fig:eval1}. Using the same coloring scheme in both figures makes it clear which of the area values in Fig.~\ref{fig:eval1} can be associated with which regions in Fig.~\ref{fig:geo1}.
To summarize, Fig.~\ref{fig:geo1}, therefore, depicts a non-static non-rectangular sum-based geometric interpretation of Eq.~\eqref{eq:ex1}. Moreover, Fig.~\ref{fig:eval1} corroborates this 
geometric interpretation.\\

\noindent
Let us know consider a different fractional integral:
\begin{equation}
\label{eq:ex2}
\frac{1}{\Gamma(\alpha)}\int_0^t f(\tau) (t-\tau)^{\alpha-1}~d\tau = \frac{1}{\Gamma(\alpha)}\int_0^t \sqrt{\tau}(t-\tau)^{\alpha-1}~d\tau. 
\end{equation}
The non-static non-rectangular sum-based geometric interpretation associated with Eq.~\eqref{eq:ex1} is depicted in Fig.~\ref{fig:geo2}. 
If we evaluate the integral in Eq.~\eqref{eq:ex1} for all $\alpha\in \mathsf{A}$ we obtain the curves in the right-most column of Table~\ref{tab:eval}. 
These curves are depicted in Fig~\ref{fig:eval2}.
The validity of the non-static non-rectangular sum-based geometric interpretation depicted in Fig.~\ref{fig:geo2} is corroborated by Fig.~\ref{fig:eval2}.

\begin{figure}[htb]
\centering
\includegraphics[width=1\textwidth]{geo_int2.pdf}
\caption{The non-static non-rectangular sum-based geometric interpretation of Eq.~\eqref{eq:ex2}.}
\label{fig:geo2}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=1\textwidth]{func_eval2.pdf}
\caption{Depicts the curves which are obtained by evaluating Eq.~\eqref{eq:ex2} for all $\alpha\in \mathsf{A}$ using either Eq.~\eqref{eq:rs} or Eq.~\eqref{eq:option2}.}
\label{fig:eval2}
\end{figure}

To conclude this section, a fractional integral of order $\alpha$ and extent $t$ can be interpreted as the area obtained by summing together the area associated with an infinite number of infinitesimals whose 
shape is determined by $\alpha$ and $t$. The geometric interpretation of a fractional integral is, therefore, non-static as it depends on the value of both $\alpha$ and $t$.

%the fractional integral in Eq.~\ref{??} can be interpreted as the area which is obtained by summing together the area associated with an infinite number of non-rectangular infinitesimals whose shape 
%is determined by the integration limits of the Cavalieri integral. Note that neither $g_t^{\alpha}(\tau)$ nor $a_t^{\alpha}(y)$ are static functions, 
%they depend on the value of $\alpha$ and $t$. This implies that a fractional integral of order $\alpha$ and extent $t$ can be interpeted as the area obtained by summing together the area associated with an infinite number of infinitesimals whose 
%shape is determined by $\alpha$ and $t$, i.e. the geometric interpretation of Eq.~\eqref{eq:ex1_raw} is non-static as it depends on the value of $\alpha$ and $t$.

% The circular markers in Figure ?? is depicted using the same tint-of-gray with which the right edge of the region  in question is depicted in Fig.
% 
% Each of the curves 
% in Fig can therefore be associated with one of the panels in Fig. The area associated with the regions depicted in each of the panels in Fig. are plotted on the curves 
% in 
% associated with a particular panel appropriate 
% curves 
% in Fig using circular markers. The circular marker is depicted in Figure using the same tint-of-gray with which the 
% right edge of the region in question is depicted in Fig.  
% 
% it is $\alpha$  
% Although $\alpha$ and $t$ alters the shape of the infinitesimals which make up the regions depicted in Fig.,  
% 
% that make up the regions in Fig are confirmed by the following observation: $\mathsf{B}_T^{\alpha}\neq\mathcal{B}_T^{\alpha}$.
% We can visually verify that this observation is in fact true for $alpha=\frac{4}{5}$ by inspecting the right-most panel in Fig.~\ref{??}. The dashed curves and the grey tinted curves in this panel do not allign.
% 
% 
% 
% the upper limit 
% of the Cavalieri integral in Eq. As the upper limit of the Cavalieri integral depends on
% 
% As the shape of the infinitesimals which make up the regions depicted in Fig depend of the integration limit of the 
% Cavalieri integral or stated differently on the value of $t$ and $\alpha$.  
% 
% If we set the value of $\alpha$ to $\frac{4}{5}$, but we do not fix the value of $t$ then Eq.~\eqref{eq:ex1} reduces to:
% \begin{equation}
% \int_{a_{t}^{\frac{4}{5}}(y)}^{b_{t}^{\frac{4}{5}}(y)} f(\tau)~d\tau 
% \end{equation}
% If we evaluate the above integral we obtain the area of the region $R_{t}^{\frac{4}{5}}$. The region $R_{t}^{\frac{4}{5}}$ is the region bounded by the $\tau$-axis, the function $f(\tau)$ and the function $b_{t}^{\frac{4}{5}}(y)$.
% Note that the shape of $R_{t}^{\frac{4}{5}}$ depends on the value of $t$. The regions contained in the set $\mathsf{R}^{\frac{4}{5}}$ are depicted in the rightmost panel of Fig\textbf{??}. Let $\mathsf{B}^{\frac{4}{5}}$ contain the right 
% edges of the regions contained in the set $\mathsf{R}^{\frac{4}{5}}$. The shape of the edges in $\mathsf{B}^{\frac{4}{5}}$ depend on the value of $t$. The edges in $\mathsf{B}^{\frac{4}{5}}$ are depicted in Fig?? using different shades of gray. The shade with which the aformentioned curves are depicted becomes lighter 
% as the value of $t$ increases. The two sets b and b are not equal, i.e. the shape of the infinitesimals that make up the regions contained in 
% R are dependent on $t$. Although the infinitesimals that make up the regions contained in R have different shapes their form do share common characteristics. 
% 
% 
% altering $t$ alters the geometric interpretaion of 
% Eq???. The shape of the inifinitesimals that make up the region that is associated with Eq depend on the value of $t$. 
% 
% 
% the shape of the integration strip one has to use to construct the infinite sum that underpins the  used to create the   . Changing the value of $t$ alters the shape of the non-rectangular integration strip which is needed 
% 
% 
% limit alters the
% shape of the infinitesimal we need to use.
% 
% 
% The shade with which the aformentioned edges are depicted becomes lighter as the value of $t$ increases.
% 
% 
% 
% The boundaries of these regions are depicted in varying shades of gray.  
% 
% 
% We depict $R_{t}^{\frac{4}{5}}$ in the left most panel of Fig.~\textbf{??} for all $t\in\mathsf{T}$. The set of functions   
% 
% 
% If we change the value of $t$ we change the area 
% 
% 
% In general can use $n$ integration strips to approximate 
% 
% 
% 
% the area If we are to sum together the area of  non-rectangular integration strips whose sides are translations of $b_{10}^{\frac{4}{5}}(y)$ then the result our summation would tend to
% the area of $R_{10}^{\frac{4}{5}}$. 
% 
% 
% Lets now choose to approximate the area of $R_{10}^{\frac{4}{5}}$ using five non-rectangualr integration strips of equal width. Mathematically speaking, this is accomplished 
% by setting $n$ to 5 in Eq??. The aformentioned integration strips are depicted in the right-most panel of Figure~\ref{??}.
% The sides of the integration strips are depicted using dashed lines. The more integration strips we consider the better our approximation of $R_{10}^{\frac{4}{5}}$ becomes. 
% 
% 
% 
% We can better approximate the area of $R_{10}^{\frac{4}{5}}$ by increasing the number of integrations strips we use .   Increasing the number of integration strips we use improves our approximation of the area of $R_{10}^{\frac{4}{5}}$  
% If we decide on using five integration strips 
% 
% 
% that 
% we use whose sides are translations of $b_{10}^{\frac{4}{5}}(y)$ 
% 
% 
% Each panel of 
% Figure~\ref{??} depicts the function $f(\tau)$ and the set of curves $\mathsf{B}_T^{\alpha}$. The set $\mathsf{B}_T^{\alpha}$ is defined in Table~\ref{??} and contains 
% all the Cavalieri integral upper limits obtained by fixing the order of integration $\alpha$ to a specific value. In each panel, we depict the set of curves $\mathsf{B}_T^{\alpha}$ using different shades of grey. 
% The shade with which $b_t^{\alpha}(y)$ is depicted becomes lighter as the value of $t$ increases.
% 
% 
% The The $y$-valued functions are depicted in different shades of gray. The $\tau$-valued fun 
% 
% The curve in the rightmost panel of 
% Figure~\ref{} which is depicted in the lightest shade of gray of 
% 
% The function $b_{10}^{\frac{4}{5}}(y)$ 
% depicted in the lightest shade of gray used within the right-most panel
% 
% 
% The right-most panel of Figure.~\ref{??} contains 
% a few $y$-valued curves, each one depicted in a different shade of gray. The curve depicted in the lightest shade of gray is  
% 
% 
% In each of the panels  curves $\mathsf{B}_1^{\alpha}$
% 
% convert the 
% integrals in ?? and ?? to equivalent Cavalieri integrals.
% 
% 
% We can assign a geometric interpretation to the above fractional integrals by converting them into Cavalieri integrals. This can be achieved by first 
% converting the above integrals into Riemann-Stieltjes integrals and then converting the resulting integrals into Cavalieri integrals. Exaclty how these conversion are 
% to be performed is discussed in the previous section. The inegrator $g_t^{\alpha}(\tau)$ of the aforementioned intermediate Riemann-Stieltjes integrals are equal and independent of the integrand $f(\tau)$.
% The set of functions $\{g_{10}^{\alpha}(\tau)|\alpha\in\{0,0.2,\cdots,0.8,1\}\}$ are presented in Table \textbf{??} and depicted in Figure\textbf{??}. The integration limits of the aformentioned Cavalieri 
% integrals are functions and are determined by the integrand $f(\tau)$ and the integrator $g_t^{\alpha}(\tau)$. The set of functions $\{a_{10}^{\alpha}(y)|\alpha\in\{0,0.2,\cdots,0.8,1\}\}$ that can be 
% associated with the integrals in Eq~{??} and Eq~{??} are presented in Table~\textbf{??}. Note that neither $g_t^{\alpha}(\tau)$ nor $a_t^{\alpha}(y)$ are static functions, 
% they depend on the value of $\alpha$ and $t$. This implies that a fractional integral of order $\alpha$ and extent $t$ can be interpeted as an infitesimal sum of an infinite number of infinitesimals whose 
% shape is determined by $\alpha$ and $t$.
% 
% Lets clarify the above by working through a detailed example. Lets set the value of $\alpha$ to $\frac{4}{5}$ and the value of $t$ to 10 and then convert the 
% integrals in ?? and ?? to equivalent Cavalieri integrals. If we evaluate these Cavalieri integrals we obtain the area of two regions $A$ and $B$. Both $A$ and $B$ are bounded by the $x$-axis, 
% $f(x)$ and $b_{10}^{\frac{4}{5}}(y)$. Region $A$ and $B$ are respectively depicted in the rightmost panel of Fig ?? and Fig ??. The area of $A$ and $B$ can be approximated
% by summing together the area of non-rectangular integration strips that inscribes $A$ and $B$. The case in which we use five integration strips is depicted in 
% Fig ?? and Fig ??. The analytic expressions of the right sides of these integration strips are elements of the set $\mathsf{A} = \{b_{10}^{\frac{4}{5}}(y) - c|c\in\{0,2,4,6,8\}\}$. The 
% curves are plotted using dashed lines. 
% 
% The curves $\mathsf{B} = \{b_{t}^{\frac{4}{5}}(y)|t\in\{2,4,6,8,10\}\}$ are plotted in different shades of grey. The shade with which $b_t^{\alpha}(y)$ is depicted becomes lighter as the value of $t$ increases. Note 
% that $\mathsf{A}\neq \mathsf{B}$. The sides of the integration strips 
% 
% 
% is the lighter the curve associated with $b_t^{\alpha}(y)$ is.
% 
% We depict We depict $b_10^{\alpha}(y)$ with the lightest
% shade of
% depicted in the right most   depict the region associated that is associated with the Cavalieri integral which is obtained 
% 
% 
% the obtained Lets depict the geometric interpretation

%\begin{equation}
%\frac{1}{\Gamma(\alpha)}\int_0^t f(\tau) (t-\tau)^{\alpha-1}~d\tau = \frac{1}{\Gamma(\alpha)}\int_0^t \sqrt{\tau}(t-\tau)^{\alpha-1}~d\tau.
%\end{equation}

%The integration limits of the aformentioned Cavalieri 
%integrals are functions and are determined by the integrand $f(\tau)$ and the integrator $g_t^{\alpha}(\tau)$. Recall that a Cavalieri integral can 
%be interpreted as an infinite sum of non-rectanguar infinitesimals whose shape is determined by its lower integration limit $a_t^{\alpha}(y)$. The set of functions 
%$\{a_{10}^{f,\alpha}(y)|\alpha\in\{0,0.2,\cdots,0.8,1\}\}$ that can be associated with the integrals in Eq~{??} and Eq~{??} are presented in Table~\textbf{??}. Note that 
%$a_t^{\alpha}(y)$ is not a static function it depends on the order of integration $\alpha$ and the integration limit $t$. The fact that $a_t^{\alpha}(y)$ is dependend 
%on $t$ implies that a fractional integral of a particular order cannot be converted to a single Cavalieri integral, but rather to an infinite set $\mathcal{C}$ of Cavalieri integrals.
%Each element of $\mathcal{C}$ is associated with a particular choice of $t$. Let $\mathcal{B}_f = \{b_{t}^{f,\alpha}(y)|\alpha\in\{0,0.2,\cdots,0.8,1\}\}$. In Fig.~\ref{fig:geo1} and Fig.~\ref{fig:geo2} we plot the elements of $\mathcal{B}$. 
%The brigther the function the larger $t$ is. For the case $t=10$ we also plot the integration strips. The fractional integral $\frac{1}{\Gamma(\alpha)}\int_0^10 f(\tau) (t-\tau)^{\alpha-1}~d\tau$ can therefore 
%be interpreted as the sum of inThey clearly do not match the other functions.

% \begin{enumerate}
%  \item Examples.
%  \item If we evaluate \textbf{[is this word correct?]} these integrals we obtaine the functions in Table \textbf{??}. These functions are plotted in Fig xx.
%  \item Assign a geometric interpretation by converting to a Riemann-Stieltjes integral and then to a Cavalieri integral.
%  \item \textbf{[Is there a need to mention the fact that it is mapped to a set of different Cavalieri integrals]}. Note that we obtain a different 
%  Cavalieri integral for each value of alpha and t.
%  \item Let us fix alpha and t to specific values, ie write down the integrals. The function b the left most function in the leftmost panel (plotted in the lightest 
%  shade of gray of all the y-valued functions. The integral , and by implication , can be inerpreted as the are of the region formed by f, b and the x-axis. This 
%  area can be obtained by summing together an infinite number of infinitesimals whose shape is determined by b. This are can be approximated by using integration strips. These 
%  strips is depicted using dashed lines. The area of this region and $t$ forms a specific point on the curves in Fig. The point is plotted in the same shade of gray than b.
%  \item If we take another value of t we obtain a different region. Note that b and b are not mere translations of each other. Changing the integration limit alters the 
%  shape of the infinitesimal we need to use. Althoug both t and alpa alters the shape of b, it is only alpha that fundamentally changes .... \textbf{[how to state this shape altering thing i do not know]} 
%  \item Need to finish the discussion of al the points mapping to the other figures. 
%  \item Summary of overall interpretation of fractional integral.....
% \end{enumerate}

% These steps 
% were discussed in more detail in the previous section. 
% 
% The steps to visualize the above two integrals were described in the previous paragraph. We first convert the integrals into Riemann-Stieltjes integrals. The 
% Riemann-Stieltjes integrals are then converted to Cavalieri integrals. We can then simply depict the Cavalieri integrals and in doing so obtain a geometric interpretaion
% of the original fractional integrals. Importantly, for each value of $t$ we obtain a different

\section{First-level section heading}

Section headings use an initial capital letter on the first word, with subsequent words lowercase.  In general, the style of the journal is to leave all section headings unnumbered.  Consult the journal editor if you wish to depart from this and other conventions.

\subsection{Second-level heading}

The same goes for second-level headings.  It is not necessary to add font commands to make the math within heads bold and sans serif; this change will occur automatically when the production style is applied.

\section{Graphics}

Figures for the \textsc{Journal} can be submitted as either color or black \& white graphics.  Generally, color graphics will be used for the online publication, and converted to black \& white images for the print journal.  We recommend using whatever graphics program you are most comfortable with, so long as the submitted graphic is provided as a separate file using a standard file format.

For best results, please follow the following guidelines:
\begin{enumerate}
\item Bitmapped file formats---preferably TIFF or JPEG, but not BMP---are appropriate for photographs, using a resolution of at least 300 dpi at the final scaled size of the image.
\item Line art will reproduce best if provided in vector form, preferably EPS.
\item Alternatively, both photographs and line art can alternatively be provided as PDF files.  Note that creating a PDF does not affect whether the graphic is a bitmap or vector; saving a scanned piece of line art as PDF does not convert it to scalable line art.
\item If you generating graphics using a \TeX\ package, please be sure to provide a PDF of the manuscript.  In the production process, \TeX-generated graphics will eventually be converted to more conventional graphics so the \textsc{Journal} can be delivered in e-reader formats.
\item For photos of contributing authors, we prefer photos that are not cropped tight to the author's profile, so that production staff can crop the head shot to an equal height and width.  If possible, avoid photographs that have excess shadows or glare.
\end{enumerate}

\section{Theorems, definitions, proofs, and all that}

Following the defaults of the \texttt{amsthm} package, styling is provided for \texttt{theorem}, \texttt{definition}, and \texttt{remark} styles, although the latter two use the same styling.

\begin{theorem}[Pythagorean Theorem]
Theorems, lemmas, axioms, and the like are stylized using italicized text. These environments can be numbered or unnumbered, at the author's discretion.
\end{theorem}

\begin{proof}
Proofs set in roman (upright) text, and conclude with an ``end of proof'' (q.e.d.) symbol that is set automatically when you end the proof environment.  When the proof ends with an equation or other non-text element, you need to add \verb~\qedhere~ to the element to set the end of proof symbol; see the \texttt{amsthm} package documentation for more details.
\end{proof}

\begin{definition}[Secant Line]
Definitions, remarks, and notation are stylized as roman text.  They are typically unnumbered, but there are no hard-and-fast rules about numbering.
\end{definition}

\begin{remark}
Remarks stylize the same as definitions.
\end{remark}

Note that the \textit{College Mathematics Journal\/} is meant to be accessible to a broad audience, so heavy use of theorem-like formalisms is generally discouraged.

\begin{acknowledgment}
The authors wish to thank the Greek polymath Anonymous, whose prolific works are an endless source of inspiration.
\end{acknowledgment}

\begin{abstract}
An abstract should not contain concrete mathematics, but rather should be discrete.  Be brief and avoid using mathematical notation except where absolutely necessary, since this brief synopsis will be used by search engines to identify your article!
\end{abstract}

\begin{thebibliography}{1}
\bibitem{parker13} Adam Parker, Who solved the Bernoulli equation and how did they do it? \textit{Coll. Math. J.} \textbf{44} (2013) 89--97.

\bibitem{hopkins} Brian Hopkins, ed., \textit{Resources for Teaching Discrete Mathematics}, Mathematical Association of America, Washington DC, 2009.
\end{thebibliography}
\vfill\eject

\end{document}