### 2019-08-19 Cav2D/article-template.tex
added some references

\documentclass{article}
\usepackage{college-math-j}
\usepackage{xfrac}
\usepackage{textcomp}

%MAKING THE CAVALIERI INTEGRAL NOTATION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\makeatletter
%\let\DOTSI\relax % amsmath support for \dots
\newcommand*{\cint}{%
  \DOTSI
  \mathop{%
    \mathpalette\@LetterOnInt{\mathbf{c}}%
  }%
  \mkern-\thinmuskip % thin space is inserted between two \mathop
  \int
}
\newcommand*{\@LetterOnInt}[2]{%
  \sbox0{$#1\int\m@th$}%
  \sbox2{$%
    \ifx#1\displaystyle
      \textstyle
    \else
      \scriptscriptstyle
    \fi
    #2%
  \m@th$}%
  \dimen@=.4\dimexpr\ht0+\dp0\relax
  \ifdim\dimexpr\ht2+\dp2\relax>\dimen@
    \sbox2{\resizebox*{!}{\dimen@}{\unhcopy2}}%
  \fi
  \dimen@=\wd0 %
  \ifdim\wd2>\dimen@
    \dimen@=\wd2 %
  \fi
  \rlap{\hbox to \dimen@{\hfil
    $#1\vcenter{\copy2}\m@th$%
  \hfil}}%
  \ifdim\dimen@>\wd0 %
    \kern.5\dimexpr\dimen@-\wd0\relax
  \fi
}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% IF YOU HAVE FONTS INSTALLED
%\usepackage{mtpro}
%\usepackage{mathtime}

\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}

\theoremstyle{definition}
\newtheorem*{definition}{Definition}
\newtheorem*{remark}{Remark}

\begin{document}

\title{Computing volume via oblique or irregularly shaped rectangular infinitesimals}
%\author{Woodrow Wilson and Herbert Hoover} %%leave blank in initial submission to allow for double blind reviewing
%\author{ }

\maketitle

%Notes authors should leave bios out altogether.

%\begin{biog} %comment out in initial submission to allow for double blind reviewing
%\item[\biogpic{\includegraphics[width=84pt]{Woodrow_Wilson.pdf}}Woodrow Wilson] (twwilson@princeton.edu) received his PhD in history and political science from Johns Hopkins University. He held visiting positions at Cornell and Wesleyan before joining the faculty at Princeton, where he was eventually appointed president of the university.  Among his proudest accomplishments was the abolition of eating clubs at Princeton on the grounds that they were elitist.

%\item[\biogpic{\includegraphics[width=84pt]{Herbert_Hoover.pdf}}Herbert Hoover] (hchoover@stanford.edu) entered Stanford University in 1891, after failing all of the entrance exams except mathematics.  He received his BS degree in geology in 1895, spent time as a mining engineer, then was appointed by his co-author to the U.S. Food Administration and the Supreme Economic Council, where he orchestrated the greatest famine relief efforts of all time.
%\end{biog}
%This is not to say that there  This is partly, because there is no simplistic geometric interpretation for fractional calculus.  
\noindent
Bonaventura Cavalieri was a mathematician who lived in the 17th century. He made numerous contributions to the fields of optics, astronomy and geometry.
He served as a professor of mathematics at the university of Bologna from 1629 until his death in 1647 \cite{andersen1985,eves1991,jullien2015}. 
Cavalieri was a student of Galileo. Galileo held Cavalieri in high esteem. Galileo made the following remark in a reference letter he wrote on Cavalieri's behalf: ``\emph{few, perhaps none, from 
Archimedes onwards, have investigated so deeply and have entered so far into the science of geometry}'' \cite{gonzalez2018}. In 1935, a moon crater, Cavalerius, was officially named after Cavalieri to commemorate his achievements and to honor his memory (see Figure~\ref{fig:crater}) \cite{blagg1935}. 
\begin{figure}[htb]
\centering
\includegraphics[width=0.4\textwidth]{crater.jpg}
\caption{The moon crater Cavalerius, which is named after Bonaventura Cavalieri. James Stuby / Wikimedia Commons / Public Domain.}
\label{fig:crater}
\end{figure}
Cavalieri's most notable achievement was the publication of his 
treatise, \emph{Geometria indivisibilibus}, in which he describes the precalculus method of indivisibles. The method of indivisibles can be attributed to the ancient Greeks,
in this case Democritus and Archimedes \cite{eves1991}. The method is used to compute the area and the volume of complex planar figures and solids, respectively.\\

\noindent
Two planar figures or solids are Cavalieri congruent if they can be decomposed into the same set of parallel ``indivisibles''. An indivisible of 
a planar figure is any cord of the planar figure. An indivisible of a solid is any planar section of a solid. Two indivisibles are equal if they are of the same length or area.
It is Cavalieri's principle that assures us that two planar figures or solids that are composed of the same set of indivisibles have equal area or volume \cite{andersen1985,eves1991,jullien2015,malik1984,wildberger2002,william1917,young1985}. Cavalieri's principle 
in $\mathbb{R}^3$ is stated below without proof:

\begin{theorem}[Cavalieri's principle in $\mathbb{R}^3$]
If two solids are included between a pair of parallel planes, and if the areas of
the two sections cut by them on any plane parallel to the including planes are always
equal, then the volumes of the two solids are equal (see Figure~\ref{fig:cav_principle}).
\end{theorem}
\noindent
The method of indivisibles works by finding a planar figure or a solid, whose area or volume can easily be computed, which is Cavalieri congruent to a planar figure or solid whose 
area or volume is difficult to compute.

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{cav_principle.pdf}
\caption{Illustration of Cavalier's principle in $\mathbb{R}^3$. The volume of the left rectangular prism is equal to the volume of the right irregularly shaped rectangular prism 
due to Cavalieri's principle in $\mathbb{R}^3$. The infinitesimals associated with these two solids are also depicted.}
\label{fig:cav_principle}
\end{figure}

\noindent
Figure~\ref{fig:cav_principle} depicts a classic rectangular prism or column (left solid) as well as an \emph{irregular rectangular prism or column} (right solid). The volumes of the two solids in Figure~\ref{fig:cav_principle} are equal due 
to Cavalieri's principle. Note that in this article we use the non-standardized phrase irregular rectangular prism or column to refer to a solid that has rectangular 
basis and lateral edges that are non-linear curves. Moreover, the four lateral edges of the aformentioned solid also need to be translations of one another. In this paper, we will also take the liberty to refer to an irregular rectangular prism whose base's
lenght and widht both tend to zero as an \emph{irregular rectangular infinitesimal}. \\

\noindent
Ackermann et al. utilized the idea of using non-rectangular integration strips to perform integration in $\mathbb{R}^2$ and in doing so developed the Cavalieri integral \cite{ackermann12}. 
The area of a non-rectangular integration strip is computed via Cavalieri’s principle. The Cavalieri integral has proven to be very useful: it can be used to visualize the Riemann-Stieltjes integral \cite{grobler19}. In this paper, we extend the Cavalieri integral into $\mathbb{R}^3$, i.e. we use Cavalieri's principle in $\mathbb{R}^3$ to define the double 
Cavalieri integral. We also show that if a solid can be decomposed into an infinite number of irregularly shaped rectangular infinitesimals then its volume can be computed 
by the double Cavalieri integral. Whether the double Cavalieri integral will prove to be an equally useful visualization tool will be investigated as part of a future endeavor.
Moreover, the examples we provide at the end of the paper are presented in such a way that they will deepen undergraduate Calculus students' understanding into the 
process of integration. Furhtermore, as this paper is mainly meant to be read by undergraduate students as well as the layman who is interested in the theory of integration, we will as far as possible try to steer clear from using too many formal definitions and proofs and rather convey our ideas via worked out examples.\\

\noindent
We start the paper by defining the single and double Cavalieri integral. We end our paper with two worked out examples.

\section{Cavalieri Integral}
The single Cavalieri integral was originally defined by Ackermann et al. without having to resort to vector calculus \cite{ackermann12,stewart1999}. In this section, however, we will define the Cavalieri integral using vector calculus notation. As the reader will undoubtedly realize, if we define the single Cavalieri integral using vector calculus then it becomes trivial to
extend its definition into $\mathbb{R}^3$.\\

%In this section we define the Cavalieri integral \cite{ackermann12}.

\noindent
Consider the continuous real-valued function $f(x)$, the continuous curve $\mathbf{c}(t)=\left <c_x(t),c_y(t) \right >$ and the interval $[a,b]$.

\begin{definition}[Translational curve in $\mathbb{R}^2$]
If a curve $\mathbf{c}(t)$ in $\mathbb{R}^2$ intersects the $x$ axis at the origin and $\mathbf{c}(t) + \left < e,0 \right >$ intersects 
the function $f(x)$ at a single point for all $e\in [a,b]$ then $\mathbf{c}(t)$ is a translational curve 
with respect to $f(x)$ on $[a,b]$. 
\end{definition}

\noindent
Moreover, assume that $\mathbf{c}(t)$ is a translational curve with respect to $f(x)$ on $[a,b]$. Let us first partition $[a,b]$ into $n$ subintervals $[x_{i},x_{i+1}]$ of equal 
width $\Delta x^1 = \Delta x_i^1 = x_{i+1}-x_i\ll (b-a)$. Furthermore, let the transformation function $h$ map the $x$--coordinate $x_i^1\in [a,b]$ at which $\mathbf{c}(t) + \left <x_i^1,0 \right >$ intersects the 
$x$-axis to the $x$--coordinate at which $\mathbf{c}(t) + \left <x_i^1,0 \right >$ intersects $f(x)$. Note, the domain of $h$ is $[a,b]$. Let $a'$ denote the $x$-coordinate 
at which $\mathbf{c}(t) + \left <a,0 \right >$ intersects $f(x)$ and let $b'$ denote the $x$-coordinate 
at which $\mathbf{c}(t) + \left <b,0 \right >$ intersects $f(x)$.  Note, the range 
of $h$ is $[a',b']$. As the partition of $[a,b]$ can be made arbitrarily fine, $h:[a,b]\rightarrow [a',b']$ actually maps every point $x_i^1\in [a,b]$ to a point $x_i^2\in [a',b']$. \\

\noindent
Let $R$ denote the region which is bounded by $f(x)$, $\mathbf{c}(t)+\left <a,0\right >$, $\mathbf{c}(t)+\left <b,0\right >$ and the $x$-axis. 
As an example, the region $R$ depicted in Figure~\ref{fig:caval2} is obtained if we set $f(x) = x$, $\mathbf{c}(t) = \frac{1}{2}\left<- t, t\right>$, $a = 1$ and $b = 4$.
In general, let $A_R$ denote the area of $R$. We can approximate $A_R$ with:
\begin{equation}
 \label{eq:caval1}
 A_R \approx \sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1.
 \end{equation}
Let us verify that this is indeed the case. If we closely inspect Equation~\ref{eq:caval1} and Figure~\ref{fig:caval2} we notice the following:
 \begin{itemize}
  \item $\Delta x_i^1 = x_{i+1}^1 - x_i^1$ represents the width of a non-rectangular integration strip.
  \item $f(x_i^2)$ represents the height of a non-rectangular integration strip.
  \item $f(x_i^2)\Delta x_i^1$ represents the area of a non-rectangular integration strip (due to Cavalieri's principle).
  \item The Cavalieri sum $\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1$, therefore, approximates the area of $R$.
  \item In the limit the Cavalieri sum (if this limit exists), therefore, approaches the area of $R$.
 \end{itemize}

\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{fig13.eps}
\caption{Region bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$, and $b(y)=4-y$. The parametric representation of $a(y)$ is $\mathbf{c}(t) = \frac{1}{2} \left <-t,t \right >$. The figure also depicts the partition points $x_i^1$ and $x_i^2$ as used in the Cavalieri sum (see Equation~\eqref{eq:caval1}). Reproduced from Quaestiones Mathematicae (2012) 35: 265-296 with permission \copyright~ NISC (Pty) Ltd.}
\label{fig:caval2}
\end{figure}

\noindent
The aforementioned limit is known as the Cavalieri integral and in this paper we denote this integral as follows:

\begin{equation}
\label{eq:cav_1d}
\int \limits_{\mathcal{C}} f(x)~dx = \lim_{n \rightarrow \infty}  \sum_{i=0}^{n-1} f(x_i^2) \Delta x_i^1,
\end{equation}
with $\mathcal{C} = \{\mathbf{c}(t),[a,b]\}$. Equation~\eqref{eq:cav_1d} is difficult to evaluate in its current form. Luckily, the Cavalieri integral 
can be converted into an equivalent Riemann integral using the transformation $h$:
 \begin{equation}
 \int \limits_{\mathcal{C}} f(x)\, dx = \int_{a}^{b} f\circ h(x)~dx. 
\end{equation}

\noindent
The only thing that now remains is to find an expression for $h$. Note that, for every coordinate $(x_i^1,x_i^2) \in [a,b] \times [a',b']$ there exists a unique $t\in\mathbb{R}$ such that
\begin{equation}
 \label{eq:pat_rel_1d}
 \left <x_i^2,f(x_i^2) \right > = \left <x_i^1,0 \right> + \left<c_x(t),c_y(t) \right >.
 \end{equation}
 We can rewrite Equation~\eqref{eq:pat_rel_1d} as
 \begin{equation}
 \label{eq:t_func_1d}
 c_y(t) = f(x_i^1 + c_x(t)).
 \end{equation}
 We can now solve for $t$ in Equation~\eqref{eq:t_func_1d} in terms of $x_i^1$. If we substitute the solution $t(x_i)$ so obtained into Equation~\ref{eq:pat_rel_1d} we obtain 
\begin{equation}
\label{eq:h_func_1d}
 x_i^2 = h(x_i^1) = x_i^1 + c_x\circ t(x_i^1).
\end{equation}

\noindent
The area of the region $R$ depicted in Figure~\ref{fig:caval2} can, therefore, be computed as follows:
\begin{equation}
\int \limits_{\mathcal{C}} f(x)~dx = \int_a^b f \circ h (x)\, dx = \dfrac{1}{2}\int_1^4x\, dx = 3.75.  
\end{equation}
The expression for $h$ was found by employing Equation~\eqref{eq:pat_rel_1d}, Equation~\eqref{eq:t_func_1d} and Equation~\eqref{eq:h_func_1d}. Let us now verify that 
this result is correct by using a more conventional approach:
\begin{equation}
\int_0^2x\, dx+\int_2^44-x\, dx- \int_0^{\frac{1}{2}}x\, dx-\int_{\frac{1}{2}}^11-x\, dx = 3.75. 
\end{equation}

\noindent
In \cite{ackermann12}, the Cavalieri integral is denoted via $\int_{a(y)}^{b(y)} f(x)~dx$, where $a(y)$ denotes the functional representation of $\mathbf{c}(t) + \left <a,0 \right >$ and $b(y) = a(y)+(b-a)$

%Let the region $R$ be bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$ and $b(y)=4-y$. This region is shown in Fig.~\ref{fig:caval2}.\\
%\begin{figure}[htb]
%\centering
%\includegraphics[width=0.3\textwidth]{fig12}
%\caption{Region bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$, and $b(y)=4-y$. Reproduced from Quaestiones Mathematicae (2012) 35: 265-296 with permission \copyright~ NISC (Pty) Ltd.}
%\label{fig:ex1}
%\end{figure}

% \noindent
% We can compute the area of $R$ as follows:
% \begin{equation}
% \int_0^2x\, dx+\int_2^44-x\, dx- \int_0^{\frac{1}{2}}x\, dx-\int_{\frac{1}{2}}^11-x\, dx = 3.75. 
% \end{equation}
% 
% \noindent
% Alternatively, the area of $R$ can also be determined via the Cavalieri integral:
% \begin{equation}
% \label{eq:caval1}
% \int_{a(y)}^{b(y)}f(x)\, dx = \lim_{n\to \infty}\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1,
% \end{equation}
% with $\Delta x_i^1 = x_{i+1}^1 - x_i^1$. Furthermore, the partition points $(x_i^1)_{i=0}^{n}$ and $(x_i^2)_{i=0}^{n}$ are depicted in Fig.~\ref{fig:caval2}. Let us review why this is in fact the case:
% \begin{itemize}
%  \item $\Delta x_i^1 = x_{i+1}^1 - x_i^1$ represents the width of a non-rectangular integration strip.
%  \item $f(x_i^2)$ represents the height of a non-rectangular integration strip.
%  \item $f(x_i^2)\Delta x_i^1$ represents the area of a non-rectangular integration strip (due to Cavalieri's principle).
%  \item The Cavalieri sum $\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1$, therefore, approximates the area of $R$.
%  \item In the limit the Cavalieri sum, therefore, approaches the area of $R$.
% \end{itemize}
% 
% \noindent
% The Cavalieri integral is, however, hard to evaluate directly. Fortunately, it is easy to convert a Cavalieri integral into an equivalent Riemann integral:
% \begin{equation}
% \int_{a(y)}^{b(y)}f(x)\, dx = \int_{a}^{b} f\circ h(x)~dx, 
% \end{equation}
% with $a(0) = a$ and $b(0)=b$. The function $h$ is known as the transformation function and it maps $x_i^1\in[a,b]$ to $x_i^2\in[a',b']$. It can be constructed via the function $g(x) = x - a\circ f(x) + a$ as 
% $h=g^{-1}$.
% 
% \noindent
% The area of $R$ can, therefore, also be computed as follows:
% \begin{equation}
% \int_{a(y)}^{b(y)}f(x)\, dx = \int_a^b f \circ h (x)\, dx = \dfrac{1}{2}\int_1^4x\, dx = 3.75.  
% \end{equation}
% 
% Let $\mathbf{c}(t) = \left <c_x(t),c_y(t) \right>$ be the vector equation of $a(y)-a$. The notation which we have been using to denote a Cavalieri integral up until now 
% in the paper was first proposed by Ackermann. Going forward in this paper, however, we propose making use of the following alternative notation:
% \begin{equation}
% \label{eq:new_def_cavalieri}
% \int \limits_{\mathcal{C}} f(x)~dx := \int_{a(y)}^{b(y)}f(x)~dx,
% \end{equation}
% with $\mathcal{C} = \{\mathbf{c}(t),[a,b]\}$. Note that $\left<a,0\right> + \mathbf{c}(t)$ and $\left<b,0\right> + \mathbf{c}(t)$ are the vector equations of $a(y)$ and 
% $b(y)$. The reason for this change in notation will become clear to the reader at a later stage. In turns out, that we can use $\mathbf{c}(t)$ to construct $h(x)$ via:
% \begin{equation}
% \label{eq:h_func}
% x_i^2 = h(x_i^1) = x_i^1 + c_x\circ t(x_i^1).
% \end{equation}
% 
% \noindent
% We will now show that this is indeed the case. Note that, for every coordinate $(x_i^1,x_i^2) \in [a,b] \times [a',b']$ there exists a unique $t\in\mathbb{R}$ such that
% \begin{equation}
% \label{eq:pat_rel}
% \left <x_i^2,f(x_i^2) \right > = \left <x_i^1,0 \right> + \left<c_x(t),c_y(t) \right >.
% \end{equation}
% We can rewrite Equation~\eqref{eq:pat_rel} as
% \begin{equation}
% \label{eq:t_func}
% c_y(t) = f(x_i^1 + c_x(t)).
% \end{equation}
% We can now solve for $t$ in Equation~\eqref{eq:t_func} in terms of $x_i^1$. If we substitute the solution $t(x_i)$ so obtained into Equation~\ref{eq:pat_rel} we obtain Equation~\ref{eq:h_func}.
% 
% In the case of $R$ we have that $\mathbf{c}(t) = \left<-\frac{1}{2} t,\frac{1}{2} t\right>$. If we solve for $t$ in terms of $x_i^1$ after having substituted $\mathbf{c}(t)$ into Eq we obtain 
% $t(x_i^1) = x_i^1$. If we now substitute our solution $t(x_i^1)$ into Eq we obtain, as required, $h(x_i^1) = \frac{1}{2}x_i^2$. This result, therefore, validates Eq ??.

\section{Double Cavalieri Integral}
In this section we define the double Cavalieri integral and we show that it can be used to calculate the volume of solids that can be decomposed into an infinite 
number of irregularly shaped regtangular infinitesimals.\\

\noindent
Without the loss of generality, consider the continuous surface $f(x,y)$, the continuous curve $\mathbf{c}(t)=\left <c_x(t),c_y(t), c_z(t) \right >$ and the region $R = [a,b]\times [c,d] = \{(x,y)\in \mathbb{R}^2|a \leq x \leq b, c \leq y \leq d\}$.

\begin{definition}[Translational curve in $\mathbb{R}^3$]
If a curve $\mathbf{c}(t)$ in $\mathbb{R}^3$ intersects the $xy$ plane at the origin and $\mathbf{c}(t) + \left < e,f,0 \right >$ intersects 
the surface $f(x,y)$ at a single point for all $(e,f)\in R$ then $\mathbf{c}(t)$ is a translational curve 
with respect to $f(x,y)$ on $R$. 
\end{definition}

\noindent
Moreover, assume that $\mathbf{c}(t)$ is a translational curve with respect to $f(x,y)$ on $R$. Let us first partition $R$ into $n\times m$ subrectangles $R_{ij} = [x_{i},x_{i+1}] \times [y_{j},y_{j+1}]$ of equal width $\Delta x^1 = \Delta x_i^1 = x_{i+1}-x_i\ll (b-a)$ and length $\Delta y^1= \Delta y_j^1 = y_{j+1}-y_j\ll(d-c)$.
Furthermore, let the transformation $T$ map the $xy$--coordinate $(x_i^1,y_j^1)\in R$ (the lower left vertice of $R_{ij}$) at which $\mathbf{c}(t) + \left <x_i^1,y_j^1,0 \right >$ intersects the 
$xy$-plane to the $xy$--coordinate at which $\mathbf{c}(t) + \left <x_i^1,y_j^1,0 \right >$ intersects $f(x,y)$. Note, the domain of $T$ is $R$. Moreover, let us denote the range 
of $T$ with $S$. As the partition of $R$ can be made arbitrarily fine, $T:R\rightarrow S$ actually maps every point $(x_i^1,y_j^1)\in R$ to a point $(x_i^2,y_j^2)\in S$. In general, the transformation $T$ is defined as follows:
\begin{equation}
x_i^2 = h(x_i^1,y_j^1)~\textrm{and}~y_j^2 = s(x_i^1,y_j^1).
\end{equation}

%We will now show 
%that we can approximate $V_{\mathcal{S}}$ by summing together the volume of $n$ non-uniform prisms that cicumscribe $\mathcal{S}$ and that as $n\rightarrow \infty$ 
%the aformentioned sum approaches $V_{\mathcal{S}}$. 
%Without any loss of generality, assume that $f(x,y)\geq0$. Also, let $S_f = \{(x,y,z)\in\mathbb{R}^3|(x,y)\in S ~ \textrm{and} ~ z=f(x,y)\}$. 
%Let $\mathcal{S}$ denote the solid which is obtained by connecting the vertices of $R$ and $S_f$ via translations of $\mathbf{c}(t)$ along 
%the $xy$ plane.

%We can construct 
%a solid We can use $\mathbf{c}(t)$, 
%$R$ and $f(x,y)$ to construct a solid $\mathcal{S}$. The bottom and the top base of $\mathcal{S}$ is $R$ and $S_f$. The solid $\mathcal{S}$ also has the following four 
%lateral edges: 
%\begin{itemize}
%\item The segement of $\mathbf{c}(t) + \left <a, c,0\right>$ which starts at $(a,c,0)$ and ends at\\$(h(a,c),s(a,c),f(h(a,c),s(a,c))$.
%\item The segement of $\mathbf{c}(t) + \left <a, d,0\right>$ which starts at $(a,d,0)$ and ends at\\$(h(a,d),s(a,d),f(h(a,d),s(a,d))$.
%\item The segement of $\mathbf{c}(t) + \left <b, c,0\right>$ which starts at $(b,c,0)$ and ends at\\$(h(b,c),s(b,c),f(h(b,c),s(b,c))$.
%\item The segement of $\mathbf{c}(t) + \left <b, d,0\right>$ which starts at $(b,d,0)$ and ends at\\$(h(b,d),s(b,d),f(h(b,d),s(b,d))$.
%\end{itemize}

\noindent
Also, let $S_f = \{(x,y,z)\in\mathbb{R}^3|(x,y)\in S ~ \textrm{and} ~ z=f(x,y)\}$.  Let us further assume that $f(x,y) > 0$ for all $(x,y)\in S$. Let $\mathcal{S}$ denote the solid which is obtained by connecting the vertices of $R$ and 
$S_f$ via translations of $\mathbf{c}(t)$ along the $xy$ plane and then closing the resulting open faces that form due to the aforementioned connections by employing appropriate surfaces. Note that 
$\mathcal{S}$ can be decomposed into an infinite number of irregularly shaped rectangular infinitesimals. As an example, the solid $\mathcal{S}$ depicted in Figure~\ref{fig:prismatoid_solid} is obtained if we set $f(x,y) = -x - y + 8$, $\mathbf{c}(t) = -\left<\frac{1}{2}t,t,t\right>$ and $R = \{(x,y)\in\mathbb{R}^2|0 \geq x \geq 4, 0 \geq x \geq 4\} \nonumber$.
Moreover, the domain $R$ and the range $S$ of the transformation function $T$ which is associated with the solid $\mathcal{S}$ is depicted in Figure~\ref{fig:prismatoid_regions}.  
In general, let $V_{\mathcal{S}}$ denote the volume of $\mathcal{S}$. We can approximate $V_{\mathcal{S}}$ with:
\begin{equation}
\label{eq:cav_sum_2d}
V_{\mathcal{S}} \approx \sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta x_i^1 \Delta y_j^1.
\end{equation}
Let us verify that this is indeed the case. If we closely inspect Equation~\ref{eq:cav_sum_2d} and Figure~\ref{fig:prismatoid_solid} we notice the following:
\begin{itemize}
\item $\Delta x_i^1$ and $\Delta y_j^1$ represent the width and the length of subrectangle $R_{ij}$.
\item The product $\Delta A = \Delta x_i^1\times \Delta y_j^1$ represents the area of subrectangle $R_{ij}$.
\item $f(x_i^2,y_j^2)$ represents the height of an irregular rectangular column or prism $\mathcal{K}_{ij}$ (or in the case of Figure~\ref{fig:prismatoid_solid} and oblique rectangular prism) whose base is $R_{ij}$. Furthermore, the lateral edges of this column are all translations 
of $\mathbf{c}(t)$ along the $xy$-plane. The cross-section between any parallel plane to the $xy$-plane and $\mathcal{K}_{ij}$ is a rectangle with the same dimensions as $R_{ij}$. 
\item The product $f(x_i^2,y_j^2)\times \Delta A$ represents the volume of $\mathcal{K}_{ij}$ (due to Cavalieri's principle).
\item The double Cavalieri sum $\sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta A$, therefore, approximates $V_{\mathcal{S}}$. 
\item In the limit the double Cavalieri sum approaches $V_{\mathcal{S}}$ (if this limit exists). 
\end{itemize}

\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{prismatoid_solid.pdf}
\caption{The solid $\mathcal{S}$ which is bounded by the planes $f(x,y)=-x-y+8$, $2x$, $2(x-4)$, $y$, $y-4$ and the $xy$-plane. The figure also depicts the partition 
points $(x_i^1,y_i^1)\in\mathcal{R}$ and $(x_i^1,y_i^1)\in\mathcal{R}$ which is used in the double Cavalieri sum (see Equation~\eqref{eq:cav_2d}).}
\label{fig:prismatoid_solid}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{prismatoid_regions.pdf}
\caption{The transformation function $T$ which is associated with the solid $\mathcal{S}$ depicted in Figure~\ref{fig:prismatoid_solid}.}
\label{fig:prismatoid_regions}
\end{figure}

\noindent
In this paper, we will refer to the aforementioned limit as the double Cavalieri integral and we will denote this integral as follows:

\begin{equation}
\label{eq:cav_2d}
\iint \limits_{\mathcal{C}} f(x,y)~dA = \lim_{n,m \rightarrow \infty}  \sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta x_i^1 \Delta y_j^1,
\end{equation}
with $\mathcal{C} = \{\mathbf{c}(t),R\}$. As is the case for the single Cavalieri integral, Equation~\eqref{eq:cav_2d} is difficult to evaluate in its current form. Luckily, the double 
Cavalieri integral can be converted into an equivalent double Riemann integral using the transformation $T$:

\begin{eqnarray}
\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \lim_{n,m\rightarrow \infty} \sum_{j=1}^n\sum_{i=1}^n f(x_i^2,y_j^2) \Delta x_i^1\Delta y_j^1\\
&=&  \lim_{n,m\rightarrow \infty} \sum_{j=1}^n\sum_{i=1}^n f(h(x_i^1,y_j^1),s(x_i^1,y_j^1)) \Delta x_i^1\Delta y_j^1\\
&=& \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy.
\end{eqnarray}

\noindent
The only thing that now remains is to find expressions for $h$ and $s$. Note that, for every coordinate pair $(x_i^1,y_j^1)\in R$ and $(x_i^2,y_j^2)\in S$ there exists a unique $t\in \mathbb{R}$ such that 
\begin{equation}
\label{eq:vector_2}
\left< x_i^2, y_j^2, f(x_i^2,y_j^2) \right > = \left < x_i^1, y_j^1, 0 \right > + \left <c_x(t),c_y(t),c_z(t)\right >  
\end{equation}
The above equation can be re-expressed as:
\begin{equation}
\label{eq:c_z}
c_z(t) = f(x_i^1 + c_x(t),y_j^1 + c_y(t)). 
\end{equation}
We can now solve for $t$ in terms of $x_i^1$ and $y_j^1$. If we substitute this solution into Equation~\ref{eq:vector_2} we obtain:
\begin{equation}
\label{eq:h_2d}
x_i^2 = h(x_i^1,y_j^1) = x_i^1 + c_x\circ t(x_i^1,y_j^1) 
\end{equation}
and 
\begin{equation}
\label{eq:s_2d}
y_j^2 = s(x_i^1,y_j^1) = y_j^1 + c_y\circ t(x_i^1,y_j^1). 
\end{equation}

\section{Prismatoid}
Let $\mathcal{S}$ be the solid which is bounded by the planes $f(x,y)=-x-y+8$, $2x$, $2(x-4)$, $y$, $y-4$ and the $xy$-plane. Solid $\mathcal{S}$ is depicted 
in Figure~\ref{fig:prismatoid_solid}, Figure~\ref{fig:prismatoid} and Figure~\ref{fig:prismatoid_real}. Solid $\mathcal{S}$ is a prismatoid, since all of its 
vertices lie in two parallel planes. Furthermore, all of its lateral faces are trapezoids or triangles (if viewed from the top) \cite{bradley1979}. Also note the following: the base of $\mathcal{S}$ is the region 
 $R = [0,4]\times [0,4]$ (if viewed from the front) and that the line of intersection between $2x$ and $y$ is $\mathbf{c}(t) = -\left<\frac{1}{2}t,t,t\right>$. The region $R$ and the line $\mathbf{c}(t)$ are
also depicted and labelled in Figure~\ref{fig:prismatoid_solid}.\\

\begin{figure}[htb]
\centering
\includegraphics[width=0.5\textwidth]{prismatoid_real.jpg}
\caption{A physical not-to-scale model of the solid $\mathcal{S}$ depicted in Figure~\ref{fig:prismatoid_solid}.}
\label{fig:prismatoid_real}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{prismatoid.pdf}
\caption{Another depiction of the solid $\mathcal{S}$ found in Figure~\ref{fig:prismatoid_solid} which highlights the fact that $\mathcal{S}$ is a prismatoid. We can, therefore, calculate its volume with the prismatoid volume equation: $V_{\mathcal{S}}= \frac{h}{6}(A_1+4A_2+A_3)$.
Note, the area of the two basis of $\mathcal{S}$ are denoted by $A1$ and $A3$, the cross-sectional area of the intersection of $\mathcal{S}$ with a plane midway between the two basis is denoted via $A2$, 
and the height of $\mathcal{S}$ is denoted by $h$. Moreover, also note that the surfaces associated with $A1$, $A2$ and $A3$ are depicted in the above figure using shaded sheets. For $\mathcal{S}$, 
$A1=10.7331$, $A2= 7.1554$, $A3 = 3.5777$ and $h = 3.5777$. The volume of $\mathcal{S}$ is, therefore, $25.6$.}
\label{fig:prismatoid}
\end{figure}

\noindent
As we alluded to in the previous section we can calculate $V_{\mathcal{S}}$ using the double Cavalieri integral:
\begin{equation}
\label{eq:cav_ex1}
V_{\mathcal{S}} = \iint\limits_{\!\mathcal{C}} f(x,y) dA =  \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy,
\end{equation}
with $\mathcal{C}=\{\mathbf{c}(t),R\}$. We can, however, only evaluate the above integral if the transformation $T$ is known (see Figure~\ref{fig:prismatoid_regions}). We can find $T$ using Equation~\eqref{eq:h_2d} and 
Equation~\eqref{eq:s_2d}. To evaluate Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} we first need to express $t$ in terms of $x_i^1$ and $y_j^1$. If we substitute $f(x,y)$ and $\mathbf{c}(t)$ into Equation~\eqref{eq:c_z} and then solve for $t$ we obtain:
\begin{equation}
t(x_i^1,y_j^1) = \frac{1}{5}x_i^1 + \frac{1}{5}y_j^1 + \frac{8}{5}. 
\end{equation}
Substituting $t(x_i^1,y_j^1)$ into Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} results in:
\begin{equation}
x_{i}^2 = h(x_{i}^1,y_{j}^1) = \frac{4}{5} x_{i}^1 - \frac{1}{5} y_{j}^1 + \frac{8}{5}  
\end{equation}
and
\begin{equation}
y_{j}^2 = r(x_{i}^1,y_{j}^1) = \frac{3}{5} x_{i}^1 - \frac{2}{5} y_{j}^1 + \frac{16}{5}.  
\end{equation}
We can now evaluate Equation~\eqref{eq:cav_ex1}:
\begin{eqnarray}
\label{eq:final_int_prismatoid}
V_{\mathcal{S}}=\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \int_0^4\int_0^4 f(h(x,y),r(x,y))~dydx\\
&=& \int_0^4 \int_0^4 -\frac{7}{5}x + \frac{3}{5} y + \frac{16}{5}~dydx\\
&=& 25.6
\end{eqnarray}

\noindent
We can easily verify that the $V_{\mathcal{S}} = 25.6$ is in fact correct, since $\mathcal{S}$ is a prismatoid. Indeed, according to the prismatoid volume 
equation,  $V_{\mathcal{S}}=25.6$. The details of this computation is described in the caption of Figure~\ref{fig:prismatoid}.

% Let $\mathbf{c}(t) = -\left<\frac{1}{2}t,t,t\right>$ be the line of intersection between the planes $2x$ and $y$. Furthermore, note that the line 
% $\mathbf{c}(t)$ passes through the origin. The line $\mathbf{c}(t)$ is depicted and labelled in Figure~\ref{fig:prismatoid_solid}. Let 
% \begin{equation}
% R = \{(x,y)\in\mathbb{R}^2|0 \geq x \geq 4, 0 \geq x \geq 4\} \nonumber
% \end{equation}
% and
% \begin{equation}
% S = \{(x,y)\in\mathbb{R}^2|-\sfrac{1}{3}y+\sfrac{8}{3}\leq x\leq -\sfrac{1}{3}y+\sfrac{16}{3}, \sfrac{1}{2} x + 4\leq y \leq \sfrac{1}{2} x + 6 \}. \nonumber
% \end{equation}
% The region $S$ and $R$ are depicted in Figure~\ref{fig:prismatoid_solid} and Figure~\ref{fig:prismatoid_regions}. Let us now subdivide $R$ into $n\times m$
% subrectangles $R_{ij} = [x_{i},x_{i+1}] \times [y_{j},y_{j+1}]$ of equal width $\Delta x^1 = \Delta x_i^1 = x_{i+1}-x_i<<(b-a)$ and length $\Delta y^1= \Delta y_j^1 = y_{j+1}-y_j<<(d-c)$. 
% Furthermore, let the transformation $T$ map the $xy$--coordinate $(x_i^1,y_j^1)\in R$ (the lower left vertice of $R_{ij}$) at which $\mathbf{c}(t) + \left <x_i^1,y_j^1,0 \right >$ intersects the 
% $xy$-plane to the $xy$--coordinate $(x_i^2,y_j^2)\in S$ at which $\mathbf{c}(t) + \left <x_i^1,y_j^1,0 \right >$ intersects $f(x,y)$ (Figure~\ref{fig:prismatoid_solid} and Figure~\ref{fig:prismatoid_regions}). 
% %The transformation $T$ is defined by the following equations:
% %\begin{equation}
% %x_i^2 = h(x_i^1,y_j^1)
% %\end{equation}
% %and
% %\begin{equation}
% %y_j^2 = s(x_i^1,y_j^1)
% %\end{equation}
% 
% %Let $S_f = \{(x,y,z)\in\mathbb{R}^3|(x,y)\in S ~ \textrm{and} ~ z=f(x,y)\}$. Note that in $\mathbb{R}^3$, $R$ has the following vertices $\mathbf{v}_{R} = \{(x,y,z)\in\mathbb{R}^3|x\in\{a,b\},y\in\{c,d\}~\textrm{and}~z=0\}$,
% %while $S_f$ has the following vertices $\mathbf{v}_{S_f} = \{(x,y,z)\in\mathbb{R}^3|x\in\{a,b\},y\in\{c,d\}~\textrm{and}~z=f(h(x,y),s(x,y))\}$. 
% 
% We can easily compute $V_{\mathcal{S}}$ using the 
% volume equation of a prismatoid. The details of this computation is described in the caption of Figure~\ref{fig:prismatoid}. According to the prismatoid volume 
% equation $V_{\mathcal{S}}=25.6$.\\

% $R = \{(x,y)\in\mathbb{R}^2|0 \geq x \geq 4, 0 \geq x \geq 4\}$
% and $S = \{(x,y)\in\mathbb{R}^2|-\sfrac{1}{3}y+\sfrac{8}{3}\geq x\leq -\sfrac{1}{3}y+\sfrac{16}{3} \}$.
% 
% 
% 
% The line $\mathbf{c}(t)$  is depicted and labelled in

% \begin{figure}[htb]
% \centering
% \includegraphics[width=0.75\textwidth]{prismatoid_xy.pdf}
% \caption{xx}
% \label{fig:prismatoid_xy}
% \end{figure}

\section{Another example}
Let $\mathcal{S}$ be the solid which is bounded by the planes $f(x,y)=\sqrt{-x^2-y^2+4}$, $\sqrt{x}$, $\sqrt{x-1}$, $\sqrt{y}$, $\sqrt{y-1}$ and the $xy$-plane. Solid $\mathcal{S}$ is depicted 
in Figure~\ref{fig:sphere_solid}. Also note the following: the base of $\mathcal{S}$ is the region 
$R = [0,1]\times [0,1]$ (if viewed from the front) and that the curve of intersection between $\sqrt{x}$ and $\sqrt{y}$ is $\mathbf{c}(t) = \left < t,t,\sqrt{t}\right >$. \\

\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{sphere_solid.pdf}
\caption{The solid $\mathcal{S}$ which is bounded by the planes $f(x,y)=\sqrt{-x^2-y^2+4}$, $\sqrt{x}$, $\sqrt{x-1}$, $\sqrt{y}$, $\sqrt{y-1}$ and the $xy$-plane.}
\label{fig:sphere_solid}
\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{sphere_regions.pdf}
\caption{The transformation $T$ associated with the solid $\mathcal{S}$ depicted in Figure~\ref{fig:sphere_solid}.}
\label{fig:sphere_regions}
\end{figure}

\noindent
Since $\mathcal{S}$ can be decomposed into an infinite number of irregularly shaped rectangular infinitesimals we can compute its volume using:
\begin{equation}
\label{eq:cav_ex2}
V_{\mathcal{S}} = \iint\limits_{\!\mathcal{C}} f(x,y) dA =  \int_a^b\int_c^d f(h(x,y),s(x,y))~dx~dy,
\end{equation}
with $\mathcal{C}=\{\mathbf{c}(t),R\}$. We, however, first need to calculate $T$ if we wish to evaluate Equation~\ref{eq:cav_ex2} (see Figure~\ref{fig:sphere_regions}). We can determine $T$ via 
Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d}. Equation~\eqref{eq:h_2d} and Equation~\eqref{eq:s_2d} in turn requires $t(x_i^1,y_j^2)$, which can be obtained via Equation~\eqref{eq:c_z}.
If we substite $f(x,y)$ and $\mathbf{c}(t)$ into Equation~\eqref{eq:c_z} and then solve for $t$ we obtain:
\begin{equation}
t = \frac{1}{4}[-2 x_{i}^1 - 2 y_{j}^1  -  1 + K(x_{i}^1,y_{j}^1)],
\end{equation}
where
\begin{equation}
K(x_{i}^1,y_{j}^1) = \sqrt{-4(x_{i}^1)^2+4x_{i}^1+8x_{i}^1y_{j}^1-4(y_{j}^1)^2+4y_{j}^1 + 33}. 
\end{equation}
As mentioned earlier, we can use $t(x_i^1,y_j^1)$ to find expressions for $h$ and $s$. The expressions of $h$ and $s$ so obtained can then be used to evaluate Equation~\eqref{eq:cav_ex2}:  
\begin{eqnarray}
\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \int_0^1\int_0^1 f(h(x,y),r(x,y))~dydx\\
&=& \int_0^1 \int_0^1 \sqrt{-(x+t(x,y))^2 - (y+t(x,y))^2 + 4} ~dydx\\
&=& 0.864043
\end{eqnarray}

\noindent
We can verify that $V_{\mathcal{S}}=0.864043$ is indeed correct via a more traditional route. We first need to project $\mathcal{S}$ onto the $xy$ plane (see Figure~\ref{fig:sphere_xy}). We can then subdivide this projection into multiple regions (see Table~\ref{tab:table1}). Doing so actually results in $\mathcal{S}$ being decomposed into multiple solids. We can then 
compute the volume of the aforementioned solids. These computed volumes can then be added together to find the total volume of $\mathcal{S}$. The calculation details 
of this approach is presented in Table~\ref{tab:table1}. Table~\ref{tab:table1} verifies that the volume of $\mathcal{S}$ is indeed $0.864043$.

\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{sphere_xy.pdf}
\caption{The projection of the solid $\mathcal{S}$ depicted in Figure~\ref{fig:sphere_solid} onto the $xy$-plane.}
\label{fig:sphere_xy}
\end{figure}

\begin{table}

\caption{Computing the volume of the solid $\mathcal{S}$ depicted in Figure~\ref{fig:sphere_solid} using a more traditional approach. The region $R_k$ is bounded by $x=a$, $x=b$, $g_1(x)$ and $g_2(x)$ (see Figure~\ref{fig:sphere_xy}). Moreover, $c_1 \approx 1.1861407$, $c_2 \approx 1.6861407$, $c_3 \approx 0.6861407$ and $c_4 \approx 1.3507811$. $r_1(x) = \frac{1}{2}[\sqrt{17-4x^2}-1]$,
$r_2(x) = \sqrt{-x^2-x+4}$, $r_3(x) = \sqrt{-x^2-(x-1)+4}$ and $r_4(x) = \frac{1}{2}[\sqrt{21-4x^2}-1]$. If we sum the last column we obtain the volume of $\mathcal{S}$ which is $0.864043$.} 
\label{tab:table1}
\begin{center}

\begin{tabular}{|l l l l l l l l|}
  $k$ & Vertices & $a$ & $b$  & $g_1$ & $g_2$ & $f_k$ & $\iint\limits_{\!R_k} f_k~dA$\\
  1 & ABD & 0 & 1& 0 & $x$ & $\sqrt{y}$ & 0.2667\\
  2 & BLED& 1& $c_1$ & $x-1$ & $x$ & $\sqrt{y} - \sqrt{x-1}$& 0.0844\\
  3 & LFE& $c_1$ & $c_2$ & $x-1$ & $r_1(x)$&$\sqrt{y} - \sqrt{x-1}$&0.0626\\
  4 & ADC& 0 & 1 & $x$ & 1 & $\sqrt{x}$ &0.2667\\
  5 & CIG& 0 & $c_3$ & 1 & $x+1$ & $\sqrt{x} - \sqrt{y-1}$ &0.0520\\
  6 & IDJG& $c_3$ & 1 & 1 & $r_2(x)$ & $\sqrt{x} - \sqrt{y-1}$&0.0716\\
  7 & DEJ& 1& $c_1$& $x$ & $r_2(x)$ & $\sqrt{x} - \sqrt{y-1}$&0.0234\\ 
  8 & EKG& $c_3$& $c_1$& $r_2(x)$ & $r_4(x)$ & $\sqrt{-x^2-y^2+4}-\sqrt{y-1}$&0.0129\\ 
  9 & EHK& $c_1$& $c_4$& $x$& $r_4(x)$ & $\sqrt{-x^2-y^2+4}-\sqrt{y-1}$&0.0055\\
  10& EMH& $c_1$& $c_4$& $r_1(x)$& $x$ & $\sqrt{-x^2-y^2+4}-\sqrt{x-1}$&0.0092\\
  11& MFH& $c_4$& $c_2$& $r_1(x)$& $r_3(x)$ & $\sqrt{-x^2-y^2+4}-\sqrt{x-1}$&0.0091\\ 
  %1 & AEM & 0 & \sfrac{8}{5}  & 0 & $2x$ &  $y$ & 2.7307 \\
  %2 & MELB & \sfrac{8}{5} & 4  & 0 & -$\frac{1}{2}x+4$ & $y$ & 8.2560 \\
  %3 & BLF & 4 & \sfrac{24}{5} & $2x-8$ & -$\frac{1}{2}x+4$ & $y-2(x-4)$ & 0.5333 \\
  %4 & ACIJ & 0 & \sfrac{4}{3} & $2x$ & 4 & $2x$ & 3.9506\\
  %5 & JIE & \sfrac{4}{3} & \sfrac{8}{5} & $2x$ & $-3x+8$ & $2x$ & 0.5058\\ 
  %6 & EIK & \sfrac{4}{3} & \sfrac{8}{5} & $-3x+8$ & 4 &  $-x-y+8$ & 0.2939 \\ 
  %7 & EKDL & \sfrac{8}{5} & 4 & $-\frac{1}{2}x+4$ & 4 &  $-x-y+8$ & 5.9520 \\ 
  %8 & LDF & 4 & \sfrac{24}{5} & $-\frac{1}{2}x+4$ & $-3x+16$ & $-x-y+8-2(x-4)$ & 0.5333 \\ 
  %9 & CGN & 0 & \sfrac{4}{5} & 4 & $2x+4$ & $2x-(y-4)$ & 0.3413 \\ 
  %10 & NCI & \sfrac{4}{5} & \sfrac{4}{3} & 4 & $-3x+8$ & $2x-(y-4)$ & 0.6068 \\
  %11 & ICH & \sfrac{4}{5} & \sfrac{4}{3} & $-3x+8$ & $-\frac{1}{2}x+6$ & $-x-y+8-(y-4)$&0.3160\\
  %12 & IHD & \sfrac{4}{3}& 4  & 4 & $-\frac{1}{2}x+6$ & $-x-y+8-(y-4)$&1.5803\\
\end{tabular}

\end{center}

\end{table}

%in In turns out that $h$ can be constructed 
%as the reader progresses through the paper. As will become apparent later on in this paper defining your Cavalieri integral in terms of $\mathbf{c}(t)$ instead of 
%$a(y)$ makes it trivial to extend its defintion into $\mathbb{R}^3$.
%In Equation~\eqref{eq:new_def_cavalieri}, we denote the $\{\mathbf{c}(t),[a,b]\}$ via $\mathcal{C}$. Furthermore, $\mathbf{c}(t)$ denotes the vector equation of 
%$a(y)-a$.
%an alternative notation

% \noindent
% Intersestingly, the definition in Eq.~\eqref{??} can also be formulated via vector calculus notation. As will become apparent later on in the article extending the definition of the Cavalieri integral as it appears in ?? into $\mathbb{R}^3$ is problematic. However,
% if the definition in ?? is first re-expressed in vector calculus notation then it becomes trivial to extend it into $\mathbb{R}^3$.\\
% 
% \noindent
% We will now show that this is indeed the case. Note that, for every coordinate $(x_i^1,x_i^2) \in [a,b] \times [a',b']$ there exists a unique $t\in\mathbb{R}$ such that
% \begin{equation}
% \label{eq:pat_rel}
% \left <x_i^2,f(x_i^2) \right > = \left <x_i^1,0 \right> + \left<c_x(t),c_y(t) \right >.
% \end{equation}
% We can rewrite Equation~\eqref{eq:pat_rel} as
% \begin{equation}
% \label{eq:t_func}
% c_y(t) = f(x_i^1 + c_x(t)).
% \end{equation}
% We can now solve for $t$ in Equation~\eqref{eq:t_func} in terms of $x_i^1$. If we substitute the solution $t(x_i)$ into Equation~\ref{eq:pat_rel} we obtain
% \begin{equation}
% \label{eq:h_func}
% x_i^2 = x_i^1 + c_x\circ t(x_i^1).
% \end{equation}
% Note that Equation~\eqref{eq:h_func} is nothing more than the function $h$ expressed in terms of $t$, i.e. $x_i^2 = h(x_i^1) = x_i^1 + c_x\circ t(x_i^1)$.
% We can, therefore compute $h(x)$ from $\mathbf{c}(t)$ and as such we can can define the Cavalieri integral in terms of $\mathbf{c}(t)$ instead of $a(y)$. In light 
% of this observation we propose the following alternative notation to denote a Cavalieri integral 
% \begin{equation}
% \int  
% \end{equation}

% and as such $t(x_i^1)$ can be used to determine $h(x_i^1)$.
% i.e. $t$ is a function 
% It is therefore possible to define $t$ in terms of $x_i^1$.  
% \noindent
% Expressing \eqref{eq:caval1} in vector calculus notation
% \begin{equation}
% \int_{a(y)}^{b(y)}f(x)\, dx =\cint_a^b f(x)~dx, 
% \end{equation}
% with $\mathbf{c}(t) = <c_x(t),c_y(t)>$ denoting a vector equation that passes through the origin and is a translation of $a(y)$. We can express $a(y)$ as $\mathbf{c}(t) + <a,0>$.
% Note that partition $x_i^1$ and $x_i^2$ are related to each other via: 
% \begin{equation}
% \label{eq:pat_rel}
% <x_i^2,f(x_i^2)> = <x_i^1,0> + <c_x(t),c_y(t)> 
% \end{equation}
% It follows from Eq.~\eqref{eq:pat_rel} that 
% \begin{equation}
% c_y(t) = f(x_i^1 + c_x(t)). 
% \end{equation}
% We can therefore express $t$ as a function of $x_i^1$. We can now compute $h(x_i^1) = x_i^1 + t(x_i^1)$.

%One first divides a complex planar figure whose area is unknown into indivisible parallel chords. These cords are then translated 
%in such a way that the original planar figure is transformed into a planar figure whose area is easily computable. The volume of a complex solid can be computed 
%in a similar manner. 
%In this method areas of complex planar figures are computed by dividing such figures into indivisible parallel chords. One then 
%finds another planar figure whose area is known that consists of the exact same set of chords. The area of the complex planar figure is then

\begin{abstract}
In this paper we extended the Cavalieri integral into three dimensional euclidean space, i.e. we defined the double Cavalieri integral. 
We also showed that this integral can be used to compute the volume of solids that can be decomposed into an infinite number of 
irrgularly shaped rectangular infinitesimals. We end the paper with two fully worked out examples. In the first example we compute the 
volume of a prismatoid using the double Cavalieri integral. In the second example we compute the volume of a complex irregular shaped solid which is housed within a sphere using the double Cavalieri integral.
\end{abstract}

\begin{thebibliography}{1}

\bibitem{ackermann12} Ackermann,~E.~R, Grobler,~T.~L., Kleynhans,~W., Olivier,~J.~C., Salmon,~B.~P, van Zyl,~A.~J. (2012). Cavalieri
Integration. \textit{Quaest. Math.} 35(3): 265–296.

\bibitem{andersen1985}
K.~Andersen. (1985). Cavalieri's Method of Indivisibles. \emph{Arch. Hist. Exact Sci.} 31(4): 291--367.

\bibitem{blagg1935}
M.~A Blagg, A. Blagg and K. M{\"u}ller. (1935). \emph{Named Lunar Formations}. Percy Lund, Humphries and Co. Ltd.

\bibitem{bradley1979}
A.~D.~Bradley. (1979). Prismatoid, Prismoid, Generalized Prismoid. \emph{Amer. Math. Monthly.} 86(6): 486--490.

\bibitem{eves1991}
H.~Eves. (1991). Two Surprising Theorems on Cavalieri Congruence. \emph{College Math. J.} 22(2): 118--124.

\bibitem{gonzalez2018} D.~L.~Gonz{\'a}lez, D.~L.~Rabiti, J.~H.~E~Cartwright. 2018. Bonaventura Cavalieri and Bologna. \emph{Math. Intelligencer}. 40(2): 21--29.

\bibitem{grobler19} Grobler,~T.~L. (2019). Visualization of the Riemann-Stieltjes integral. \textit{Coll. Math. J.} 50(3): 198--209.

\bibitem{jullien2015} V.~Jullien (Ed.). (2015). \emph{Seventeenth-Century Indivisibles Revisited}. Boston: Birkh{\"a}user.

\bibitem{malik1984}
H.~A~Malik. (1984). A Note on Cavalieri Integration, \emph{Math. Mag.} 57(3): 154--156.

\bibitem{stewart1999}
J. Stewart. (1999) \emph{Calculus, Early Transcendentals}. Brooks/Cole Publish Company.

\bibitem{wildberger2002}
W.~N.~Wildeberger. (2002). A New Proof of Cavalieri's Quadrature Formula. \emph{Amer. Math. Monthly.} 109(9): 843--845.

\bibitem{william1917}
E.~G.~William. (1917). Cavalieri's Theorem in his own Words, \emph{Amer. Math. Monthly.} 95(10): 447--451.

\bibitem{young1985}
L. Lay-Young, S. Kangshen. (1985). The Chinese Concept of Cavalieri's Principle and Its Application, \emph{Hist. Math.}, 12: 219--228.

\end{thebibliography}
\vfill\eject

\end{document}